---
title: "Spring 2026 Semester Report"
subtitle: "Representative Sample Willingness Survey"
format: html
bibliography: mocareferences.bib
citation-location: document
reference-location: section
editor: visual
---

```{r}
#| label: Data Fetching. Load R packages
#| include: false 

library(osfr) 
library(dplyr)
library(ez) 
library(tidyverse) 
library(gt)
library(gtsummary)
library(kableExtra)
library(stringr)
library(ggridges)
library(lubridate)  # for date handling
library(psych) # for factor analysis 
library(pheatmap) # for factor analysis heat map 
library(corrplot) # for correlation matrix 
library(broom) # for correlation tests
library(purrr) # for correlation tests 
library(likert) # for likert plots 
library(factoextra)
library(ggfortify)
library(car) 
library(GPArotation) # for factor analysis 
# for mlr
# library(GGally)
# library(car)
# library(MASS)
# library(glmnet)

```

```{r}
#| label: Prolific demographics data
#| include: false 

## add which round of the pilot it was by reading in demographics files so we know completion dates

#read in prolific demographics files 
demo_df <- read_csv("prolific_demographic_export_6978ff4366f4253e5b853496.csv") %>%
  mutate(Age = as.numeric(Age))

num_no_code_subjects_df <- nrow(demo_df %>%
  filter(`Completion code` == "NOCODE") %>% 
  group_by(`Participant id`) %>%
  summarize(num_no_code_subjects = n()))

#clean df
cleaned_prolific_demographics_df <- demo_df %>%
# make name the same as in data csvs so merging is easier
rename("prolific_subject_id" = "Participant id") %>%
# filter out people who didn't finish
filter(Status != "RETURNED") %>%
#rename "Started at" to make grepl function easier to use later on
rename("start" = "Started at") %>%
# make time more readable but keep the seconds version for analysis stuff later
mutate(`time_taken_overall_minutes` = as.numeric(`Time taken`) / 60) %>%
#remove the no code and unknown code people 
  filter(`Completion code` == "C12ZD3N0")

str(cleaned_prolific_demographics_df)

```

```{r}
#| label: Retrieve osf data
#| include: false

# loading in the data after manually downloading from osf and putting in a folder called data 
# not using osf retrieval because osf retrieval wasn't grabbing all the csvs:
file_names <- list.files("data/", pattern = "\\.csv$", full.names = FALSE)

raw_dfs_tibbles <- map(file_names, ~read_csv(file.path("data/", .)))

rawwwwwww_df <- data.table::rbindlist(raw_dfs_tibbles)
 

#note that fnxwd is the osf node for representative sample

```

```{r}
#| label: Verifying that we have all data 
#| include: false 

# verify that we have the data for all the people by making sure that the data df has as many subs as the demo df, and printing the missing id's if not 

difference <- setdiff(
  cleaned_prolific_demographics_df$prolific_subject_id,
  rawwwwwww_df$prolific_subject_id
)

if (length(difference) > 0) {
  print("MISSING DATA FROM PARTICIPANTS:")
  print(difference)
}

```

```{r}
#| label: Merge prolific demos + create a condition type column
#| include: false 

raw_df <- rawwwwwww_df %>%
  # inner join so we drop anyone from the osf data that wasn't in the demographics file (like test runs) and so we drop anyone in the prolific file who didn't make it to OSF
  inner_join(cleaned_prolific_demographics_df, by = "prolific_subject_id") %>%
  # remove unnecessary syntax from topicChoiceAsString for readability later
  mutate(topicChoiceAsString = gsub('\\{"row":"', '', topicChoiceAsString)) %>%
  mutate(topicChoiceAsString = gsub('"}', '', topicChoiceAsString)) %>%
  # fill condition column and row and topic strings
  group_by(prolific_subject_id) %>%
  fill(topicChoiceAsString, .direction = "downup") %>%
  ungroup() 

# number of subjects in data 
total_num_subjects_in_data <- nrow(raw_df %>%
  group_by(prolific_subject_id) %>%
  summarize(total_num_subjects_in_data = n()))

```

```{r}
#| label: Clean the survey df
#| include: false

surveys_cleaning1_df <- raw_df %>%
  # Filter for only surveys
  filter(trial_type == "survey") %>%
  # Remove control interventions (these are coded as surveys)
  filter(phase != "control-intervention") %>%
  # Remove the matrix question titles because they mess with separating questions from answers
  # openmindedness and unwillingness reason
  mutate(across(everything(), ~ gsub(',"openmindedness":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"openmindedness":', '', .))) %>%
  mutate(across(everything(), ~ gsub(',"unwillingness-reason":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"unwillingness-reason":', '', .))) %>%
  # republican/democrat (pre-intervention)
  mutate(across(everything(), ~ gsub(',"rating-republicans":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub(',"rating-democrats":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"rating-republicans":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"rating-democrats":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"interacting-with-republicans":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"interacting-with-democrats":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"interacting-with-republicans":null', '', .))) %>%
  mutate(across(everything(), ~ gsub('"interacting-with-democrats":null', '', .))) %>%
  mutate(across(everything(), ~ gsub('"rating-republicans":', '', .))) %>%
  mutate(across(everything(), ~ gsub('"rating-democrats":', '', .))) %>%
  mutate(across(everything(), ~ gsub('"interacting-with-republicans":', '', .))) %>%
  mutate(across(everything(), ~ gsub('"interacting-with-democrats":', '', .))) %>%
  # republican/democrat (post-intervention)
  mutate(across(everything(), ~ gsub('"rating-republicans-post":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"rating-democrats-post":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"interacting-with-republicans-post":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"interacting-with-democrats-post":null,', '', .))) %>%
  mutate(across(everything(), ~ gsub('"rating-republicans-post":', '', .))) %>%
  mutate(across(everything(), ~ gsub('"rating-democrats-post":', '', .))) %>%
  mutate(across(everything(), ~ gsub('"interacting-with-republicans-post":', '', .))) %>%
  mutate(across(everything(), ~ gsub('"interacting-with-democrats-post":', '', .))) %>%
  # the name of the question in 2.1 run of the experiment
  mutate(across(everything(), ~ gsub('topicChoice":', '', .))) %>%
  # the name of the question in future runs of the experiment
  mutate(across(everything(), ~ gsub('topic":', '', .))) %>%
  mutate(across(everything(), ~ gsub(',null,', '', .))) %>%
  # there is a bug (ish) in the slider feature in jspsych: the slider sits at 50 on the user end, but the data is entered as null unless the user moves the slider around (even if only to put it back at 50). four participants encountered this problem, so i'm going to correct their data 
  mutate(across(everything(), ~ gsub('"slider1-placeholder":null,', '"slider1-placeholder":50,', .))) %>%
  mutate(across(everything(), ~ gsub('"slider2-placeholder":null,', '"slider2-placeholder":50,', .)))


surveys_cleaning2_df <- surveys_cleaning1_df %>%
  # Remove weird characters but don't remove quotations because we need them for separating questions
  mutate(across(everything(), ~ gsub('}', '', .))) %>%
  mutate(across(everything(), ~ gsub('\\{', '', .))) %>%
  mutate(across(everything(), ~ gsub('\\[', '', .))) %>%
  mutate(across(everything(), ~ gsub('\\]', '', .))) %>%
  # Make each question its own row
  separate_rows(response, sep = ',"') %>%
  # Remove any leftover quotation marks
  mutate(across(everything(), ~ gsub('"', '', .))) %>%
  # Move answers to a new column
  separate(response, into = c("question", "answer"), sep = ":", extra = "merge", fill = "right") %>%
  # Remove all remaining questions that were answered with null (like the slider questions, because the slider 'placeholder' questions are where people actually answer the slider questions)
  filter(!is.na(answer) & answer != 'null' & answer != 'null,null') %>%
  filter(question != "unwillingness-reason-")

```

```{r}
#| label: Reverse code necessary items 
#| include: false 

# reverse code open-mindedness items 

surveys_cleaning3_df <- surveys_cleaning2_df %>%
    mutate(answer = ifelse(
     grepl("@R@", question),
     8 - as.numeric(answer),
     (answer)
   ))

```

```{r}
#| label: Remove unnecessary columns 
#| include: false 

surveys_df <- surveys_cleaning3_df %>%
  select(-c(stimulus, trial_type, trial_index, time_elapsed, internal_node_id, study_id, session_id, unwillingness_yes_rows, "Submission id", Status, "Custom study tncs accepted at", "Reviewed at", "Archived at", "Completion code", "Total approvals"))

```

```{r}
#| label: Create a df that flags people who failed the attention check or commitment check
#| include: false


#see who passes the check. I'm including people who may have left the slider at 1 because this was almost certainly just an honest mistake in not moving the slider far enough 
attention_check <- surveys_df %>%
  filter(question == "attention-check-slider-pre-placeholder") %>%
  group_by(prolific_subject_id) %>%
  mutate(n_hits = sum(answer %in% c("0.00", "1.00")),
    flag = case_when(
      n_hits == 0 ~ "fail (single)",
      n_hits  > 0 ~ "pass",
      TRUE ~ NA_character_ )) %>%
  ungroup() %>%
  select(flag, n_hits, question, answer, prolific_subject_id) 

#get list of all subject id's so we can see which subjects didn't even answer the question (the sliders aren't recorded in our df if the subject doesn't move them. they naturally sit at 50.) 
all_ids <- surveys_df %>%
  distinct(prolific_subject_id)

#filter for the question of interest
attention_ids <- surveys_df %>%
  filter(question == "attention-check-slider-pre-placeholder") %>%
  distinct(prolific_subject_id) 

#see who's missing
missing_ids <- setdiff(all_ids$prolific_subject_id, attention_ids$prolific_subject_id)

#make a table for these missing people 
missing_rows <- tibble(
  flag = "fail (single)",
  question = NA_character_,
  answer = NA_character_,
  prolific_subject_id = missing_ids,
  time_taken_overall_minutes = NA_real_
)

# combine those who answered and those who didn't answer
attention_check <- bind_rows(attention_check, missing_rows)

# extract value for text 
total_failed_attention_checks <- attention_check %>%
  # excluding people for both fails
  filter(flag == 'fail (single)') %>%
  summarize(total_failed_attention_checks = n_distinct(prolific_subject_id)) %>%
  pull(total_failed_attention_checks)

#extract list of ids so we can remove the data
failed_attention_check_ids <- attention_check %>%
  filter(flag == "fail (single)") %>%
  distinct(prolific_subject_id) %>%
  pull(prolific_subject_id)


# check the commitment check
fail_commitment_check <- surveys_df %>%
  filter(question == "commitment-check") %>%
  filter(answer != "yes")
```

```{r}
#| label: Remove subs who failed attention check 
#| include: false

# remove bad subs
raw_df <- raw_df %>%
  filter(!(prolific_subject_id %in% failed_attention_check_ids))
    
surveys_df <- surveys_df %>%
  filter(!prolific_subject_id %in% failed_attention_check_ids) 

total_num_clean_subjects_in_data <- nrow(raw_df %>%
  group_by(prolific_subject_id) %>%
  summarize(total_num_clean_subjects_in_data = n()))

```

```{r}
#| label: Print surveys_df csv (and txt for Claude)
#| echo: false 

write.csv(surveys_df, "rep_sample_survey_data.csv", row.names = FALSE)

write.table(surveys_df, file = "pretest_survey_data.txt", sep = "\t", row.names = FALSE, quote = FALSE)

```

## Introduction

Existing literature indicates that a complex interplay of psychological factors—including fear of social isolation (@NoelleNeumann1974; @Chen2018; @Haug2025), identity protection (@Tajfel1979; @Hunter1996), communication anxiety (@Beatty1987; @Schulenberg2023), and emotional dysregulation (@Masters2019; @Aldao2010)—converge with sociocultural influences to deter open dialogue (@Haug2025; @Kahneman2011; @Lewandowsky2017). This avoidance is a significant barrier to constructive discourse. Understanding this phenomenon is critical for fostering democratic citizenship and cultivating inclusive learning environments.

I developed an individual-level intervention to help increase Americans' willingness to engage in conversations on divisive topics. Before I can assess the effectiveness of this intervention, I need to identify the proper ways to measure willingness to engage; more specifically, I need to identify the proper ways of measuring what concerns an individual has about these conversations and how severely each one detracts from the individual's willingness to converse. With these metrics, I can identify which concerns my intervention ameliorates, if any.

Thus, to properly measure the barriers to engagement and assess the effectiveness of our intervention, I have been designing a questionnaire to capture the core concerns that contribute to an individual's reluctance to engage in conversations on controversial topics. In the fall 2025 semester, this questionnaire consisted of 19 items; I used an Asprey CCAS SRE award to deploy it, along with some additional qualitative and quantitative questions, to 200 participants to assess whether our survey is comprehensive and whether any underlying concern profiles exist among participants. The results suggested that there were 5 concerns that our survey did not properly capture and one existing item that should have been divided into two. I updated the survey accordingly, resulting in a 25-item questionnaire.

At the start of the Spring 2026 semester, I used DSS funding to deploy the updated survey and accompanying questions to a sample of Prolific participants that represents the U.S. population's distribution of sex, age, and ethnicity. This survey allows me to assess whether the updated survey properly captures the primary concerns that contribute to Americans' unwillingness to engage in productive, open-minded discussions about divisive topics, and whether there are any distinct, underlying factors that organize these concerns. After completing data collection, I spent the remainder of the semester analyzing the data, as well as constructing a plan for moving forward with publishing these findings and with applying them to the individual-level intervention.

Additionally, at the start of the semester I made a Vassar-specific, mobile-friendly version of the survey. This can be found at https://jodeleeuw.github.io/URSI-LLM-Convo-Bot/experiment/index_vassar.html. I designed flyers with survey-linked QR codes and distributed them around Vassar College campus with the goal of gathering data from Vassar students; however, at the time of writing this report, I only have data from five participants. This is likely due to a lack of incentives to participate and the saturation of other surveys and flyers around campus, especially because I did not post flyers in hotspot areas (e.g., bathroom stalls) per campus regulations.

The following report details the methods in collecting and analysis of the representative sample's data.

## Methods

### Materials

We used custom software written with the jsPsych framework (@jspsychcite) to create the study, which participants viewed and responded to via their personal laptops or desktops. The Javascript code for the experiment and RStudio code for data analysis can both be found at (<https://github.com/jodeleeuw/URSI-LLM-Convo-Bot/tree/main/pilot_1).> Pre-registrations and all data for this study and previous pilots are available on the Open Science Framework at <https://osf.io/x32pv.>

### Participants and Procedure

```{r}
#| label: Find time taken stats
#| echo: false 

#finding overall mean, min, max of how long participants took 
time_stats <- surveys_df %>%
  mutate(time_taken_overall_minutes = as.numeric(time_taken_overall_minutes)) %>%
  ungroup() %>%
  summarise(
    Mean = round(mean(time_taken_overall_minutes, na.rm = TRUE), 1),
    Min = round(min(time_taken_overall_minutes, na.rm = TRUE), 1),
    Max = round(max(time_taken_overall_minutes, na.rm = TRUE), 1),
    Median = round(median(time_taken_overall_minutes, na.rm = TRUE), 1)
  ) 

mean_survey_time <- time_stats$Mean
min_survey_time <- time_stats$Min
max_survey_time <- time_stats$Max
median_survey_time <- time_stats$Median
```

```{r}
#| label:  tbl-basic-demographics-table
#| echo: false
#| tbl-cap: Table of basic demographics.

#reloading package because functions from other packages were overriding??
library(gtsummary)

#we already read in demographics earlier 

#create a demographics table from prolific screening demo data which we loaded onto the raw df earlier
survey_demographics_df <- surveys_df %>%
filter(question == "commitment-check") %>% # picking random question so there's only one row per subject
select(prolific_subject_id, Age, Sex, "Ethnicity simplified", "Employment status", time_taken_overall_minutes) %>%
  mutate(Age = as.numeric(Age))

survey_demographics_df <- survey_demographics_df %>%
  mutate(across(where(is.factor), ~ droplevels(.)))

survey_demographics_df$time_taken_overall_minutes <- round(
  as.numeric(survey_demographics_df$time_taken_overall_minutes), 
  1
)

demographics_table <- tbl_summary(survey_demographics_df,
 by = NULL,
  label = list(
    `time_taken_overall_minutes` = "Time taken (minutes)",
    Age = "Age (years)",
    Sex = "Sex",
    `Ethnicity simplified` = "Ethnicity"
  ),
  statistic = NULL,
#list(
  #  all_continuous() ~ "{mean} ({sd})",
  #  all_categorical() ~ "{n} ({p}%)"
#),
  type = list(
    `time_taken_overall_minutes` = "continuous",
    `Age` = "continuous",
    `gender-identification` = "categorical",
   `Sex` = "categorical",
    `Ethnicity simplified` = "categorical"
  ),
  digits = list(
    all_continuous() ~ 2,
    all_categorical() ~ 0
  ),
 include = c(
'time_taken_overall_minutes',
'Age',
'Sex',
'Ethnicity simplified'
))
 # %>%  modify_header(label = "**whatever label i want**")

# Print the table
as_gt(demographics_table)

```

Through Prolific, 300 subjects participated in this survey. Participants were restricted to U.S. residents who fluently speak English. We used Prolific's representative sample feature to select a sample that is representative of the distributions of sex, age, and ethnicity across the U.S. Data from `r num_no_code_subjects_df` subjects were not collected because the subjects did not wait to be redirected to Prolific at the study's conclusion before exiting out of the webpage. Data from `r length(difference)` subject is missing due to a bug with DataPipe. We excluded data from `r total_failed_attention_checks` subjects for failing an attention check question, leaving us with a final sample of `r total_num_clean_subjects_in_data` subjects.

### Procedure

After answering a data integrity commitment question, participants were first asked to "Imagine [they were] about to engage in a conversation with a small group of people where differing opinions exist" and to rate on a 7-point Likert scale their willingness to engage for each of ten controversial conversation topics. These topics were: mandating vaccines in the U.S., human euthanasia in the U.S., deportation policies for undocumented immigrants in the U.S., transgender athletes in U.S. sports, the role of the U.S. government in healthcare, U.S. public policy on climate change, the use of A.I. in traditionally human-run spaces (e.g., medicine, art, driving), the role of capitalism in the U.S. economy, the role of the U.S. in the Israeli-Palestinian conflict, and police conduct and accountability in the U.S. The topic that they were least willing to discuss was selected as their focus topic for the remainder of the survey.

Next, participants were asked to rate their agreement with five statements about their open-mindedness, such as whether they avoid arguments that they disagree with. 

Participants then rated two more questions about their willingness to converse about their focus topic with individuals rather than a small group. They rated these on a slider from 0 (Absolute unwillingness) to 100 (Absolute willingness), with 50 as a neutral point. The first question asked about willingness to converse with someone who disagrees with the participant's views, and the second question asked about willingness with someone who would push against the participant's views. 

Following the overall willingness assessment, participants engaged with the 25-item concern inventory. This occurred first in a binary endorsement phase and then in a rating phase. In the binary endorsement phase, participants were asked whether each item would concern them even a little bit when having a serious conversation about their focus topic. Then, in the rating phase, participants rated the severity of each concern's effect on their willingness to converse; they rated on a 7-pt scale. Participants only saw items in the rating phase that they endorsed in the binary phase.

The session concluded with a free-response question about participants' reasons for reluctance with an emphasis on any reasons that were not captured in the concern inventory. This question asked participants to think back to the last conversation they had that they were unwilling to engage in so as to provide a memory for reflection. Participants were then debriefed, thanked for their time, and sent back to Prolific. 

## Results

### Topic-Level Willingness Rating Distributions

```{r}
#| label: Make short topic labels
#| echo: false

# shortened topic labels for use in figures
topic_labels <- c( 
  "mandating vaccines in the U.S." = 
    "Vaccine Mandates",
  "human euthanasia in the U.S." = 
    "Human euthanasia",
  "deportation policies for undocumented immigrants in the U.S." = 
    "Deportation policies",
  "transgender athletes in U.S. sports" = 
    "Transgender athletes in sports",
  "the role of the U.S. government in healthcare" = 
    "Government in healthcare",
  "U.S. public policy on climate change" = 
    "Climate change policies",
  "the use of A.I. in traditionally human-run spaces (e.g., medicine, art, driving)" = 
    "A.I. in human-run spaces",
  "the role of capitalism in the U.S. economy" = 
    "Capitalism in economy",
  "the role of the U.S. in the Israeli-Palestinian conflict" = 
    "U.S. in Israeli-Palestinian conflict",
  "police conduct and accountability in the U.S." =
    "Police conduct and accountability"
)

# really short labels. should only use if necessary 
really_short_topic_labels <- c( 
  "mandating vaccines in the U.S." = 
    "Vaccines",
  "human euthanasia in the U.S." = 
    "Euthanasia",
  "deportation policies for undocumented immigrants in the U.S." = 
    "Deportation",
  "transgender athletes in U.S. sports" = 
    "Trans athletes",
  "the role of the U.S. government in healthcare" = 
    "Healthcare",
  "U.S. public policy on climate change" = 
    "Climate change",
  "the use of A.I. in traditionally human-run spaces (e.g., medicine, art, driving)" = 
    "A.I.",
  "the role of capitalism in the U.S. economy" = 
    "Capitalism",
  "the role of the U.S. in the Israeli-Palestinian conflict" = 
    "Israel-Palestine",
  "police conduct and accountability in the U.S." =
    "Police"
)


# make into dfs and add to surveys_df for ease of use  
short_topic_labels_lookup <- tibble(
  short_topic_labels = unlist(topic_labels),
  topicChoiceAsString = rep(names(topic_labels), lengths(topic_labels))
)

really_short_topic_labels_lookup <- tibble(
  really_short_short_topic_labels = unlist(really_short_topic_labels),
  topicChoiceAsString = rep(names(really_short_topic_labels), lengths(really_short_topic_labels))
)

surveys_df <- surveys_df %>%
  merge(short_topic_labels_lookup, by = "topicChoiceAsString") %>%
  merge(really_short_topic_labels_lookup, by = "topicChoiceAsString")

```

```{r}
#| label: Make focus topic dfs and grab counts for each topic
#| echo: false

# ratings of all topics
topic_ratings_df <- surveys_df %>%
  filter(phase == "pre-convo-survey-initial") %>%
  filter(question != "commitment-check") %>%
  mutate(answer = as.numeric(answer)) %>%
  select(prolific_subject_id, question, answer, topicChoiceAsString)

# ratings of just focus topic
topic_choice_rating <- topic_ratings_df %>%
  group_by(prolific_subject_id) %>%
  filter(question == topicChoiceAsString)

# counts and add short labels for ease of use later
topic_frequency_counts <- topic_choice_rating %>%
  group_by(question) %>%
  count() %>%
  rename(topicChoiceAsString = question) %>%
  merge(short_topic_labels_lookup, by = "topicChoiceAsString")
```

```{r}
#| label: fig-topic-ratings
#| echo: false
#| fig-cap: Each subject used a 7-pt scale (where 1 meant strongly uncomfortable and 7 meant strongly comfortable) to rate how comfortable they would be discussing each of the depicted controversial topics. The topic that a subject rated lowest, meaning most uncomfortable, became the subject's focus topic for the remainder of the survey. This ridgeline plot depicts the distributions of all topic ratings. 

topic_ratings_plot <- ggplot(topic_ratings_df, aes(x = answer, y = question, fill = question)) +
  geom_density_ridges(bandwidth = 0.931) + # using the bandwidth that it automatically picks 
  # give the qualitative meanings of 1-7
  scale_x_continuous(breaks = seq(1, 7, by = 1)) +
   scale_y_discrete(labels = topic_labels) +
  # make it look nicer
  theme_ridges() +      
  # remove the legend
  theme(legend.position = "none") +
  labs(title = "Ridgeline Plot of Ratings by Topic",
       subtitle = "Includes all subjects' ratings of all ten topics",
       x = "Rating (1 = Strongly uncomfortable, 7 = Strongly comfortable)",
       y = "Contentious Topic")

print(topic_ratings_plot)
```

```{r}
#| label: fig-topic-ratings-only-chosen-topics
#| echo: false
#| fig-cap: Each subject used a 7-pt scale (where 1 meant strongly uncomfortable and 7 meant strongly comfortable) to rate how comfortable they would be discussing each of the depicted controversial topics. The topic that a subject rated lowest, meaning most uncomfortable, became the subject's focus topic for the remainder of the survey. This ridgeline plot depicts the distributions of focus topic ratings. 

# reorder the factor levels based on n so it is easier to interpret in the graph version
topic_choice_rating <- topic_choice_rating %>%
  mutate(topicChoiceAsString = factor(topicChoiceAsString, 
                                      levels = topic_frequency_counts %>% 
                                        arrange(n) %>%  
                                        pull(topicChoiceAsString)))

# make ridgeline plot with the counts of participants in each topic at the side for easy visual 
topic_choice_ratings_plot <- ggplot(topic_choice_rating, aes(x = answer, y = topicChoiceAsString, fill = topicChoiceAsString)) +
  geom_density_ridges(bandwidth = 0.931) +
  scale_x_continuous(breaks = seq(1, 7, by = 1)) +
  scale_y_discrete(labels = topic_labels) +
  theme_ridges() +      
  theme(legend.position = "none") +
  labs(title = "Ridgeline Plot of Ratings by Focus Topic",
       x = "Rating (1 = Strongly uncomfortable, 7 = Strongly comfortable)",
       subtitle = "Includes all subjects' ratings only of focus topics",
       y = "Contentious Topic") +
  geom_text(
    data = topic_frequency_counts,
    aes(
      x = Inf,
      y = as.factor(topicChoiceAsString),
      label = paste0("n = ", n)
    ),
    inherit.aes = FALSE,
    hjust = 1.1,
    size = 4
  ) +
  guides(fill = "none") +
  coord_cartesian(clip = "off")

print(topic_choice_ratings_plot)
```

```{r}
#| label: fig-bar-chart-topic-choice
#| include: false 
#| fig-cap: Bar chart showing the number of participants in each topic.

topic_votes_plot <- ggplot(topic_choice_rating, aes(x = answer, fill = topicChoiceAsString)) + 
  geom_bar() + 
  xlab("Willingness (1 = Strongly Unwilling, 7 = Strongly Willing)") +
  ylab("Number of participants")


topic_votes_plot
```

```{r}
#| label: Find number of participants who rated their focus topic 4 or greater
#| echo: false

topic_choice_rating_4_plus <- topic_choice_rating %>%
  filter(answer > 3)

n_subs_4_plus <- length(topic_choice_rating_4_plus$prolific_subject_id)

```

```{r}
#| label: fig-bar-topic-counts
#| echo: false 
#| fig-cap: Bar plot showing the number of participants in each topic.

topic_hist <- ggplot(topic_choice_rating, aes(x = topicChoiceAsString)) + 
  geom_bar() + 
  xlab("Focus topic") +
  scale_x_discrete(labels = really_short_topic_labels) +
  ylab("Number of participants")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

topic_hist
```

Our topic assignment method successfully generated a sample for whom engaging with their focus topic felt less than strongly comfortable, although not as uncomfortable as we were aiming for. Only `r round(100*(total_num_clean_subjects_in_data - n_subs_4_plus)/total_num_clean_subjects_in_data, 0)`% of subjects rated their focus topic with a rating from 1 ("Strongly uncomfortable") to 3 ("Slightly uncomfortable"), whereas `r round(100*(n_subs_4_plus/total_num_clean_subjects_in_data), 0)`% of subjects rated their focus topic with a rating from 4 ("Neutral") to 6 ("Moderately comfortable"). At a minimum, though, no participant rated their focus topic willingness at a 7 ("Strongly comfortable"), implying that no participant was wholly comfortable discussing their focus topic.

The participants with higher willingness ratings were not evenly distributed amongst the topics. As shown in @fig-topic-ratings-only-chosen-topics, four of the ten topics---police conduct and accountability, vaccine mandates, capitalism in economy, and climate change policy---proportionally garnered many more willingness ratings at 4 ("Neutral") or greater than the other topics. These topics were of the less popular topics (@fig-bar-topic-counts), but they still had a considerable number of participants, especially capitalism in the economy and climate change policies.

As shown in @fig-bar-topic-counts, the majority of participants were least willing to discuss the role of the U.S. in the Israeli-Palestinian conflict, human euthanasia in the U.S., and transgender athletes in U.S. sports. Additionally, while deportation policy was not a popular focus topic, the ratings for this topic were lower (less willing), on average, than for the other focus topics.

```{r}
#| label: extra text i removed
#| include: false

# Specifically, `r topic_frequency_counts$n[topic_frequency_counts$topicChoiceAsString == "police conduct and accountability in the U.S."]` participants were least willing to discuss police conduct and accountability but still rated themselves as slightly to strongly comfortable discussing the topic, `r topic_frequency_counts$n[topic_frequency_counts$topicChoiceAsString == "mandating vaccines in the U.S."]` participants were least willing to discuss vaccine mandates but still rated themselves as slightly to strongly comfortable discussing the topic, `r topic_frequency_counts$n[topic_frequency_counts$topicChoiceAsString == "the role of capitalism in the U.S. economy"]` participants were least willing to discuss capitalism in economy but still rated themselves as slightly to strongly comfortable discussing the topic, and topic_frequency_counts$n[topic_frequency_counts$topicChoiceAsString == "U.S. public policy on climate change"] participants were least willing to discuss climate change but still rated themselves as slightly to stronly comfortable discussing the topic.
```

```{r}
#| label: fig-avg-willingness-by-topic
#| include: false # this is such a fugly graph
#| fig-cap: Ridgeline plot of willingness to engage in conversation about topic by topic. Participants only rated their willingness for their topic choice. Participants' willingness values were calculated through averaging the responses to the questions, "Using the slider, please rate your willingness to have a conversation with someone who strongly disagrees with you about [topic], where a score of 0 is absolute unwillingness to converse and 100 is absolute willingness to converse. A score of 50 is neutral," and "Using the slider, please rate your willingness to have a conversation with someone who strongly disagrees with you about [topic], where a score of 0 is absolute unwillingness to converse and 100 is absolute willingness to converse. A score of 50 is neutral."

#make avg willingness df by taking mean of both slides for each participant 

avg_willingness_df <- surveys_df %>%
  filter(question == "slider1-placeholder" | question == "slider2-placeholder") %>% 
  group_by(prolific_subject_id, topicChoiceAsString) %>%
  summarise(avg_willingness = mean(as.numeric(answer), na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(topicChoiceAsString = factor(topicChoiceAsString, 
                                      levels = topic_frequency_counts %>% 
                                        arrange(n) %>%  # Use arrange(desc(n)) for descending
                                        pull(topicChoiceAsString)))

#merge back with surveys_df in case i want all info in one spot in the future 
surveys_df <- avg_willingness_df %>%
  select(-topicChoiceAsString) %>%
  full_join(surveys_df, by = "prolific_subject_id")

topic_choice_ratings_plot <- ggplot(avg_willingness_df, aes(x = avg_willingness, y = topicChoiceAsString, fill = topicChoiceAsString)) +
  geom_density_ridges(bandwidth = 0.931) + # using the bandwidth that it automatically picks 
  scale_x_continuous() +
   scale_y_discrete(labels = topic_labels) +
  # make it look nicer
  theme_ridges() +      
  # remove the legend
  theme(legend.position = "none") +
  labs(title = "Ridgeline Plot of Willingness to Engage by Topic",
       x = "Rating (0 = Strongly unwilling, 100 = Strongly willing)",
       y = "Contentious Topic")

topic_choice_ratings_plot

```

```{r}
#| label: fig-box-plot-willingness 
#| echo: false
#| fig-cap: Distribution of average willingness ratings by topic from 0 (completely unwilling) to 100 (completely willing), where 50 is neutral. Number of participants can be found at the right edge of the figure. We asked participants to rate their willingness to discuss their focus topic with someone who strongly disagrees with them and with someone who would push back against their views, and then averaged the two results to find a participant's average willingness rating. 

#boxplot with participant counts
avg_willingness_box_plot <- ggplot(
  avg_willingness_df,
  aes(
    x = avg_willingness,
    y = topicChoiceAsString,  
    fill = topicChoiceAsString
  )
) +
  geom_boxplot() +
  geom_text(
    data = topic_frequency_counts,
    aes(
      x = Inf,
      y = as.factor(topicChoiceAsString),
      label = paste0("n = ", n)
    ),
    inherit.aes = FALSE,
    hjust = 1.1,
    size = 4
  ) +
  labs(
    title = "Boxplots of Willingness to Engage by Topic",
    x = "Rating (0 = Strongly unwilling, 100 = Strongly willing)",
    y = "Contentious Topic"
  ) +
  scale_y_discrete(labels = topic_labels) +
  guides(fill = "none") +
  coord_cartesian(clip = "off")

avg_willingness_box_plot

```

```{r}
#| label: Find n participants who rated willingness at or above 50 
#| include: false 

avg_willingness_50_plus_df <- avg_willingness_df %>%
  filter(avg_willingness > 49)

n_avg_willingness_50_plus <- length(avg_willingness_50_plus_df$prolific_subject_id)
```

After assigning focus topics, we asked participants to rate their willingness to discuss their focus topic with someone who strongly disagrees with them and with someone who would push back against their views. Participants used a slider ranging from 0 (absolute unwillingness) to 100 (absolute willingness) to respond to each question, and we averaged the two responses to find participants' average willingness. This assessment overlapped substantially with the focus topic selection question; however, we asked these questions separately to ensure quality of data and to use alternative phrasing that emphasized the potential divisiveness of the discussion.

Participants' average willingness ratings, shown in @fig-box-plot-willingness, were widely spread, with a notable density of `r round(100*(n_avg_willingness_50_plus/total_num_clean_subjects_in_data), 0)`% of participants reporting scores at or above the neutral point of 50. While not completely in opposition with the Likert-scale indicator of willingness, these results suggest an even greater average willingness to converse than did the Likert-scale indicator.

```{r}
#| label: Diff btwn two avg willingness questions 
#| echo: false 

will_df <- surveys_df %>%
  filter(question == "slider1-placeholder" | question == "slider2-placeholder") %>%
  mutate(answer = as.numeric(answer)) %>%
  select(prolific_subject_id, question, answer) %>%
  pivot_wider(
    names_from = question,
    values_from = answer
  ) %>%
  mutate(
    diff_avg_willingness = as.numeric(`slider1-placeholder`) - as.numeric(`slider2-placeholder`)
  ) %>%
  select(prolific_subject_id, "slider1-placeholder", "slider2-placeholder", diff_avg_willingness)

diff_will_df <- will_df %>%
  filter(abs(diff_avg_willingness) > 5)

n_subs_diff_will_ratings <- length(diff_will_df$prolific_subject_id)
# there are some p big differences for a handful of people, so it's good that we have both questions

```


```{r}
#| label: fig-avg-willingness-diff
#| echo: false
#| fig-cap: Slope plot depicting each subject's ratings of the first and second willingness questions. Each line is a subject, and the slope conveys the difference in ratings. We applied jitter to the y-axis to allow for visibility of density. The first willingness question asked participants to rate their willingness to discuss their focus topic with someone who strongly disagrees with them. The second question asked about willingness to discuss with someone who would push back against their views. 

##this plot is the most informative but it might be too confusing to readers 

will_for_plot_df <- will_df %>%
  select(-diff_avg_willingness) %>%
  pivot_longer(cols = c("slider1-placeholder", "slider2-placeholder"),
               names_to = "question",
               values_to = "answer") %>%
  mutate(answer = as.numeric(answer)) %>%
  mutate(has_difference = prolific_subject_id %in% diff_will_df$prolific_subject_id)


ggplot(will_for_plot_df, 
       aes(x = question, y = answer, group = prolific_subject_id)) +
  geom_line(aes(color = has_difference), alpha = 0.3, 
            position = position_jitter(width = 0, height = 2, seed = 42)) +
  geom_point(alpha = 0.5, size = 2,
             position = position_jitter(width = 0, height = 2, seed = 42)) +
  #stat_summary(aes(group = 1), fun = mean, geom = "line", color = "red", linewidth = 1.5) + # mean line. i don't think it helps with anything really
  scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "orange"),
                     labels = c("TRUE" = "Ratings are different (> 5 units)", "FALSE" = "Ratings are the same (within 5 units)"),
                     name = NULL  # remove legend title
  ) +
  scale_x_discrete(labels = c("First Question (Disagrees)", "Second Question (Pushes Back)")) +
  scale_y_continuous(n.breaks = 10) +
  labs(title = "Willingness to Discuss: Disagreement vs. Pushback",
       x = "Willingness Question (Discussion Partner Type)",
       y = "Willingness Rating (0-100)") +
  theme_minimal() 

```

An additional curious finding is that many people gave substantially different ratings to the two questions. As seen in @fig-avg-willingness-diff, `r round(100*(n_subs_diff_will_ratings/total_num_clean_subjects_in_data), 0)`% of participants responded to the two willingness questions with ratings that differed by more than 5 units. The possibility that semantic framing would influence ratings is partly what drove us to include two questions rather than one, so this finding is not necessarily a bad sign. At the same time, though, it is somewhat surprising that the difference between discussing with someone who "disagrees with you" as opposed to someone who would "push back" is so great to many participants.

#### Brief investigation of data quality indicators

The proportion of participants who gave greatly different willingness ratings, combined with the higher-than-expected willingness ratings overall, raised a slight concern that data quality may be less than ideal, so I exploratorily analyzed the few bits of information that I have that might provide some indications of data quality.

```{r}
#| label: fig-density-time-taken-diff-will
#| fig-cap: "Density plot depicting distributions of time taken on survey by the intra-subject difference between ratings of our two willingness to converse questions. The first question asks about willingness to discuss focus topic with someone who disagrees with the subject's views. The second question asks about willingness to discuss focus topic with someone who would push back on the subject's views. This was created in an effort to explore data quality, but should be interpreted very lightly."
#| echo: false
#| message: false
#| warning: false

#attach diff_avg_willingness back to surveys_df so all data is in one df
surveys_df <- surveys_df %>%
  left_join(will_df, by = "prolific_subject_id") %>%
  rename(avg_will_disagree = "slider1-placeholder",
         avg_will_pushback = "slider2-placeholder") %>%
  mutate(diff_avg_willingness = as.numeric(diff_avg_willingness)) %>%
  mutate(time_taken_overall_minutes = as.numeric(time_taken_overall_minutes))

# look at completion times

quality_ind_willingness <- surveys_df %>%
  mutate(five_threshold = case_when(
    abs(diff_avg_willingness) > 4 ~ "different",
    .default = "same"
  )) %>%
  mutate(ten_threshold = case_when(
    abs(diff_avg_willingness) > 9 ~ "different",
    .default = "same"
  )) %>%
  mutate(interval = case_when(
    abs(diff_avg_willingness) > 4 & abs(diff_avg_willingness) < 16 ~ "five_to_fifteen",
    abs(diff_avg_willingness) > 15 & abs(diff_avg_willingness) < 26 ~ "fifteen_to_twentyfive",
    abs(diff_avg_willingness) > 25 & abs(diff_avg_willingness) < 36 ~ "twentyfive_to_thirtyfive",
    abs(diff_avg_willingness) > 35 & abs(diff_avg_willingness) < 46 ~ "thirtyfive_to_fortyfive",
    abs(diff_avg_willingness) > 45 & abs(diff_avg_willingness) < 56 ~ "fortyfive_to_fiftyfive",
    abs(diff_avg_willingness) > 55 & abs(diff_avg_willingness) < 66 ~ "fiftyfive_to_sixtyfive",
    abs(diff_avg_willingness) > 65 & abs(diff_avg_willingness) < 76 ~ "sixtyfive_to_seventyfive",
    abs(diff_avg_willingness) > 75 & abs(diff_avg_willingness) < 86 ~ "seventyfive_to_eightyfive",
    abs(diff_avg_willingness) > 85 & abs(diff_avg_willingness) <= 100 ~ "eightyfive_to_hundred",
    .default = "zero_to_five"
  )) %>% # do the same thing again but with nice formatting for figure. there is a better way to do this but this does the job 
  mutate(interval_formatted = case_when(
    abs(diff_avg_willingness) > 9 & abs(diff_avg_willingness) < 16 ~ "5-15 Unit Difference",
    abs(diff_avg_willingness) > 15 & abs(diff_avg_willingness) < 26 ~ "16-25 Unit Difference",
    abs(diff_avg_willingness) > 25 & abs(diff_avg_willingness) < 36 ~ "26-35 Unit Difference",
    abs(diff_avg_willingness) > 35 & abs(diff_avg_willingness) < 46 ~ "36-45 Unit Difference",
    abs(diff_avg_willingness) > 45 & abs(diff_avg_willingness) < 56 ~ "46-55 Unit Difference",
    abs(diff_avg_willingness) > 55 & abs(diff_avg_willingness) < 66 ~ "56-65 Unit Difference",
    abs(diff_avg_willingness) > 65 & abs(diff_avg_willingness) < 76 ~ "66-75 Unit Difference",
    abs(diff_avg_willingness) > 75 & abs(diff_avg_willingness) < 86 ~ "76-85 Unit Difference",
    abs(diff_avg_willingness) > 85 & abs(diff_avg_willingness) <= 100 ~ "86-100 Unit Difference",
    .default = "0-4 Unit Difference"
  ), 
  
  interval_formatted = factor(
    interval_formatted,
    levels = c(
      "0-4 Unit Difference",
      "5-15 Unit Difference",
      "16-25 Unit Difference",
      "26-35 Unit Difference",
      "36-45 Unit Difference",
      "46-55 Unit Difference",
      "56-65 Unit Difference",
      "66-75 Unit Difference",
      "76-85 Unit Difference",
      "86-100 Unit Difference"
    )
  )
) %>%
  distinct(prolific_subject_id, .keep_all = TRUE) %>%
  ungroup()

ggplot(quality_ind_willingness, aes(x = time_taken_overall_minutes, y = interval_formatted)) +
  geom_density_ridges() +
  labs(title = "Time Taken on Study by Difference in Willingness Q Ratings",
       x = "Time taken overall on study (minutes)",
       y = "Absolute value of rating difference (0 to 100)")


# look at number of concerns endorsed. this is also interesting just in general, so I'm gonna grab some distribution stats here to report on later 

n_concerns_endorsed <- surveys_df %>%
  filter(str_detect(question, "unwillingness-reason-")) %>%
  filter(phase == "willingness-concerns-first-survey") %>%
  group_by(prolific_subject_id) %>%
  count(answer == 1) %>%
  rename(n_endorsed = n) %>%
  mutate(n_endorsed = as.numeric(n_endorsed)) %>%
  filter(`answer == 1` == TRUE) %>%
  select(prolific_subject_id, n_endorsed)

sum_stats_concerns_endorsed <- n_concerns_endorsed %>%
  ungroup() %>%
  summarize(mean = round(mean(as.numeric(n_endorsed), na.rm = TRUE), 1),
            max = max(n_endorsed),
            min = min(n_endorsed),
            lower_quartile = quantile(n_endorsed, probs = 0.25, na.rm = TRUE),
            upper_quartile = quantile(n_endorsed, probs = 0.75, na.rm = TRUE)
  )
```

```{r}
#| label: fig-density-num-concerns-diff-will
#| fig-cap: "Density plot depicting distributions of number of willingness concerns selected by the intra-subject difference between ratings of our two willingness to converse questions. The first question asks about willingness to discuss focus topic with someone who disagrees with the subject's views. The second question asks about willingness to discuss focus topic with someone who would push back on the subject's views. This was created in an effort to explore data quality, but should be interpreted very lightly."
#| echo: false
#| message: false
#| warning: false

surveys_df <- surveys_df %>%
  left_join(n_concerns_endorsed, by = "prolific_subject_id", relationship = "many-to-many") 

quality_ind_willingness <- quality_ind_willingness %>%
  left_join(n_concerns_endorsed, by = "prolific_subject_id") 

ggplot(quality_ind_willingness, aes(x = n_endorsed, y = interval_formatted)) +
  geom_density_ridges() +
  labs(title = "Number of Concerns Endorsed by Difference in Willingness Q Ratings",
       x = "Number of concerns endorsed (1 to 25)",
       y = "Absolute value of rating difference (0 to 100)")

```

Neither @fig-density-time-taken-diff-will nor @fig-density-num-concerns-diff-will indicate any drastic red flags. The time taken by individuals with a greater difference between their two willingness question ratings does not appear to be substantially less than the time taken by individuals with little to no difference. Likewise,  the number of concerns endorsed by individuals with greater rating differences does not appear to be substantially different from individuals with little to no difference. 

```{r}
#| label: fig-sliders-density
#| message: false
#| warning: false
#| include: false
#| fig-cap: Density plot displaying subjects' self-reported willingness to converse about their topic with somebody who disagrees with them and/or will challenge their views. Subjects used a scale from 0 to 100, where a rating of 0 meant subject was absolutely unwilling and a rating of 100 meant subject was absolutely willing. Subjects were asked this question in two ways; this density plot shows the mean of the two ratings. 

sliders_df <- surveys_df %>%
  filter(question == "slider1-placeholder" | question == "slider2-placeholder") %>%
  group_by(prolific_subject_id) %>%
  summarize(mean_willingness = round(mean(as.numeric(answer), na.rm = TRUE), 1))
  
sliders_density <- ggplot(sliders_df, aes(x = mean_willingness)) + #, color = condition.x)) +
  geom_density() +  
  labs(x = "Mean Willingness Rating", y = "Density of Rating", title = "Average Willingness to Converse about Selected Topic")  #, color = "Group") +
 # scale_color_manual(values = c("control free reflection" = "darkseagreen", "selection bot" = "cornflowerblue", "combination bot" = "navyblue"), 
 #                    labels = c("control free reflection" = "Control group", "selection bot" = "Experimental group: Selection bot", "combination bot" = "Experimental group: Combination bot"))

print(sliders_density)

```

```{r}
#| label: Make various short labels for concern items 
#| echo: false

concern_map <- c("Might experience negative feelings about self" = "unwillingness-reason-The conversation could negatively affect how I feel about myself during and/or afterward (e.g.,  fear, sadness, anger, vulnerability)", 
         "Might be offended" = "unwillingness-reason-I might take offense to what someone says",
  "Opinion is unpopular amongst group" = "unwillingness-reason-I would feel like my opinion is unpopular amongst the group",
  "Don't participate in these conversations often" = "unwillingness-reason-I don’t often participate in conversations like these",
  "Might struggle to explain views or seem uninformed" = "unwillingness-reason-I might struggle to explain my views to the others and/or come across as uninformed",
  "Others might be unproductive or disrespectful" = "unwillingness-reason-I would not trust the others to keep the conversation productive and respectful (e.g., keeping emotions in check, refraining from making hostile remarks, not dominating the conversation)",
  "Views wouldn't change" = "unwillingness-reason-It would be a waste of time because my views wouldn't change",
  "Don't want to listen to others" = "unwillingness-reason-I would not want to listen to the others discuss their views",
  "Might be criticized" = "unwillingness-reason-Someone could criticize or show disapproval of my views",
  "Might experience negative feelings about the world" = "unwillingness-reason-The conversation could negatively affect how I feel about the world during and/or afterward",
  "Conversation might be awkward or tense" = "unwillingness-reason-The conversation could become awkward or tense",
  "Might feel embarrassed" = "unwillingness-reason-I could feel humiliated or embarrassed in the conversation",
  "Might offend someone" = "unwillingness-reason-Someone might take offense to what I say",
  "Might struggle to remain productive and respectful" = "unwillingness-reason-I might struggle to remain productive and respectful",
  "Others won't make effort to listen" = "unwillingness-reason-The others might not make a full effort to hear and understand me",
  "Identity might be challenged" = "unwillingness-reason-Someone could challenge ideas or beliefs that play an important role in making me who I am",
  "Might feel invalidated" = "unwillingness-reason-I might feel disempowered, unheard, or invalidated",
  "Might incur social repercussions" = "unwillingness-reason-I could incur social repercussions (e.g., being excluded by others in the future, negative changes to my reputation, feeling ostracized)",
  "Opinion is unpopular in general" = "unwillingness-reason-My opinion is unpopular in general",
  "Topic is too complex" = "unwillingness-reason-I would feel like the topic is too complex or requires more knowledge than I have",
  "No first-hand experience or relevance" = "unwillingness-reason-I don't have first-hand experience with the topic and/or it isn't relevant to me",
  "Non-social repercussions (e.g., work)" = "unwillingness-reason-I could experience significant repercussions at work or in other spaces beyond my personal social network",
  "Too mentally draining" = "unwillingness-reason-It wouldn't be worth the emotional and/or mental energy",
  "Others' views won't change" = "unwillingness-reason-It would be a waste of time because the others' views wouldn't change",
  "Might damage relationships" = "unwillingness-reason-I would not want to hurt my relationship(s) with others or upset them"
)

# make tibble using these 
concern_lookup <- tibble(
  question = unlist(concern_map),
  short_concern_label = rep(names(concern_map), lengths(concern_map))
)

# create short labels for each question
question_key_for_fa <- surveys_df %>%
  filter(phase == "follow-up-unwillingness-concerns-survey") %>%
  ungroup() %>%
  select(question) %>%
  distinct(question) %>%
  arrange(question) %>%  # ensure consistent order
  mutate(short_label = paste0("Q0", row_number())) %>%
  merge(concern_lookup, by = "question")

question_key_for_printing <- question_key_for_fa %>%
  mutate(across(everything(), ~ gsub('unwillingness-reason-', '', .))) 

```

```{r}
#| label: Most frequently endorsed concerns
#| echo: false 

percent_concerned_df <- surveys_df %>%
  filter(phase == "willingness-concerns-first-survey") %>%
    # Keep only the columns needed
    select(question, answer) %>%
    # Group by the question text
    group_by(question) %>%
    # Summarize the data for each question
    summarise(
        # Calculate the count of "Yes" answers for this question
        N_yes = sum(answer == "1", na.rm = TRUE),
        # Calculate the percentage
        percent_yes = (N_yes / n()) * 100,
        .groups = 'drop' # Ungroup after summarizing
    ) %>%
    # Select and rename the final columns
    select(
        concern = question, 
        percent_yes
    ) %>%
    # Format to 1 decimal place
    mutate(
        percent_yes = round(percent_yes, 1)
    ) %>%
    # Order the results from highest to lowest percentage
    arrange(desc(percent_yes))


# select the top 5 most frequently endorsed 
top_5_concerns <- percent_concerned_df %>%
  slice(1:5) %>%
  mutate(across(everything(), ~ gsub('unwillingness-reason-', '', .))) %>%
 pull(concern)  

```

```{r}
#| label: Most severely impacting willingness concerns 
#| echo: false

percent_severe_df <- surveys_df %>%
  filter(phase == "follow-up-unwillingness-concerns-survey") %>%
    # Keep only the columns needed
    select(question, answer) %>%
    # Group by the question text
    group_by(question) %>%
    mutate(across(everything(), ~ replace_na(.x, 1))) %>% # making the concerns that participants weren't even worried about into a 1 so they can be represented properly 
    # Summarize the data for each question
    summarise(
        # Calculate the count of "Yes" answers for this question
        N_severe = sum(answer == "7" | answer == "6", na.rm = TRUE),
        # Calculate the percentage
        percent_severe = (N_severe / n()) * 100,
        .groups = 'drop' # Ungroup after summarizing
    ) %>%
    # Select and rename the final columns
    select(
        concern = question, 
        percent_severe
    ) %>%
    # Format to 1 decimal place
    mutate(
        percent_severe = round(percent_severe, 1)
    ) %>%
    # Order the results from highest to lowest percentage
    arrange(desc(percent_severe))


# select the top 5 most frequently endorsed 
top_5_severe <- percent_severe_df %>%
  slice(1:5) %>%
  mutate(across(everything(), ~ gsub('unwillingness-reason-', '', .))) %>%
 pull(concern)  

```

```{r}
#| label: fig-openmindedness-density
#| include: false
#| message: false
#| warning: false
#| fig-cap: Scatterplot with line of equality of pre- and post-conversation mean ratings of subjects' open-mindedness when discussing their topic. Subjects rated their open-mindedness on a scale from 0 to 7, where a rating of 0 meant subject was not open-minded and a rating of 7 meant subject was open-minded.

openmindedness_df <- surveys_df %>%
  filter(grepl("openmindedness", question)) %>%
  mutate(answer = as.numeric(answer)) 

openmindedness_summary_by_subject_df <- openmindedness_df %>%
  group_by(prolific_subject_id) %>%
 summarise(
    mean_openmindedness = round(mean(answer, na.rm = TRUE), 1),
    min_openmindedness = round(min(answer, na.rm = TRUE), 1),
    max_openmindedness = round(max(answer, na.rm = TRUE), 1),
    median_openmindedness = round(median(answer, na.rm = TRUE), 1)
  )  %>%
  ungroup()

openmindedness_summary_df <- openmindedness_df %>%
  ungroup() %>%
 summarise(
    mean_openmindedness = round(mean(answer, na.rm = TRUE), 1),
    min_openmindedness = round(min(answer, na.rm = TRUE), 1),
    max_openmindedness = round(max(answer, na.rm = TRUE), 1),
    median_openmindedness = round(median(answer, na.rm = TRUE), 1)
  )  %>%
  ungroup()

mean_openmindedness <- openmindedness_summary_df$mean_openmindedness
min_openmindedness <- openmindedness_summary_df$min_openmindedness
max_openmindedness <- openmindedness_summary_df$max_openmindedness
median_openmindedness <- openmindedness_summary_df$median_openmindedness


openmindedness_density <- ggplot(openmindedness_summary_df, aes(x = mean_openmindedness)) + #, color = condition.x)) +
  geom_density() +  
  labs(x = "Mean Open-mindedness Rating", y = "Density of Rating", title = "Average Open-mindedness about Selected Topic")  #, color = "Group") +
 # scale_color_manual(values = c("control free reflection" = "darkseagreen", "selection bot" = "cornflowerblue", "combination bot" = "navyblue"), 
 #                    labels = c("control free reflection" = "Control group", "selection bot" = "Experimental group: Selection bot", "combination bot" = "Experimental group: Combination bot"))

print(openmindedness_density)

```

```{r}
#| label: Making free response csv for chat analysis 
#| echo: false

free_response_df <- surveys_df %>%
  filter(question == "willingness-free-response") %>%
  ungroup() %>%
  select(prolific_subject_id, answer, topicChoiceAsString, Age, Sex, "Ethnicity simplified", "Employment status") %>%
  rename(
    #Condition = condition,
    'Focus topic' = 'topicChoiceAsString',
    'Free response' = 'answer'
    )

write.csv(free_response_df, "rep_sample_free_response_data.csv", row.names = FALSE)
```

```{r}
#| label: Df for 28 manually reviewed free responses 
#| echo: false 

# I used random.org to randomly select 28 numbers from 1 to 284, which will correspond to the row number in the free_response df

# 1. 5
# 2. 69
# 3. 131
# 4. 86
# 5. 134
# 6. 214
# 7. 240
# 8. 42
# 9. 94
# 10. 276
# 11. 187
# 12. 29
# 13. 171
# 14. 55
# 15. 184
# 16. 49
# 17. 201
# 18. 165
# 19. 176
# 20. 14
# 21. 216
# 22. 283
# 23. 1
# 24. 103
# 25. 153
# 26. 218
# 27. 284
# 28. 252

free_response_manual_df <- free_response_df %>%
  filter(prolific_subject_id == "564dd73edc21cd000dba8ce1" |
           prolific_subject_id == "610a151c96de07b15f28c8cb" |
           prolific_subject_id == "662fecef2a008847dc4f0d9b" |
           prolific_subject_id == "631c8e97db06f601f81bd82f" |
           prolific_subject_id == "6631997e0554bb6d25062fd4" |
           prolific_subject_id == "67d6ffe7dad4354b208e186d" |
           prolific_subject_id == "694f4d4cf7f1d3401de99e61" |
           prolific_subject_id == "5e0bd25e67f1703eafedd160" |
           prolific_subject_id == "63fb73b8e64f455117b045b9" |
           prolific_subject_id == "697cd31d69677a82b9e94fb3" |
           prolific_subject_id == "673f58421b143a15bc2367c1" |
           prolific_subject_id == "5cab69f77ea4a700019a87dc" |
           prolific_subject_id == "66e9d794bd67a9a02ca66a8f" |
           prolific_subject_id == "5fc07d1d954a2251529dcfd6" |
           prolific_subject_id == "5ed559e5afbd830e2117ed55" |
           prolific_subject_id == "67364113bdafe2840fa05b69" |
           prolific_subject_id == "678041088952b9da0a4734ae" |
           prolific_subject_id == "66ccd41993bcfa83410adbf0" |
           prolific_subject_id == "670d2af9c00f086b201e9f70" |
           prolific_subject_id == "5a4ff86beedc32000141f232" |
           prolific_subject_id == "67e8aa454848956e5e32c19d" |
           prolific_subject_id == "697d4f4c2229572d1e5d26a3" |
           prolific_subject_id == "5563984afdf99b672b5749b6" |
           prolific_subject_id == "64bbf16db125b5d6a2f3a911" |
           prolific_subject_id == "66779b54dfca48b1e171a63b" |
           prolific_subject_id == "67ebc24ee98e314102e23f1a" |
           prolific_subject_id == "697d5370cf685ff023402820" |
           prolific_subject_id == "696125abb83cbbfa966d0e64"
           )

```

```{r}
#| label: tbl-free-response-examples-table
#| include: false
#| tbl-cap: Table showing examples of uncaptured concerns in subjects' responses to free response question, which asked subjects "Think back to the last time you chose to not participate in a conversation about [topic] or another topic you found uncomfortable to discuss. In the space below, please describe why you decided to not engage. Consider whether your reasons align with any of the potential concerns listed in this survey, or if you had different concerns."

# This table should show a couple very typical responses, which, based on LLM analysis, should probably be consistent with our twenty-five items 

# free_response_for_table <- free_response_df %>%
#    filter(prolific_subject_id == "5d3695969749cc00165c222f"
#           | prolific_subject_id == "66293ab7b323f5f7ce785c38"
#           | prolific_subject_id == "60712d937752fb8780e89951"
#           | prolific_subject_id == "61074ae762b0186a302f1028"
#           | prolific_subject_id == "565bff58c121fe0005fc390d"
#           | prolific_subject_id == "666efed91a31f3c82021f11d") %>%
#    select("Free response", "Focus topic")

# kable(free_response_for_table, "html") %>%
#   kable_styling() %>%
#   column_spec(2, width = "5em")
```

### Distributions of Concerns Ratings & Comprehensiveness

```{r}
#| label: fig-unwillingness-reason-likert-bar-plot
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Likert bar plot displaying subjects' ratings of how much certain concerns contribute to their unwillingness to converse. The 25 potential concerns were derived from existing literature and pretested in a pilot study. First, participants indicated whether or not each item has any effect at all on their willingness. Then, for the concerns that do, they rated the severity of the effect on a scale from 1 to 7. A rating of 1 meant a participant's willingness is not at all impacted by the concern and a rating of 7 meant a participant's willingness is strongly impacted by the concern. For this figure, when a participant did not endorse a concern in the binary phase as having any effect on their willingness, I coded the severity of the concern's effect on their willingness as a 1.

# only works though if all "1" to "7" ratings are in the df

wide_unwill_reasons_df_1 <- surveys_df %>%
  filter(grepl("unwillingness-reason", question)) %>%
  filter(phase == "follow-up-unwillingness-concerns-survey") %>%
  select(question, answer, prolific_subject_id) %>%
  group_by(prolific_subject_id) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(id_cols = prolific_subject_id, names_from = question, values_from = answer) %>%
  ungroup() %>%
  rename("Might experience negative feelings about self" = "unwillingness-reason-The conversation could negatively affect how I feel about myself during and/or afterward (e.g.,  fear, sadness, anger, vulnerability)", 
         "Might be offended" = "unwillingness-reason-I might take offense to what someone says",
  "Opinion is unpopular amongst group" = "unwillingness-reason-I would feel like my opinion is unpopular amongst the group",
  "Don't participate in these conversations often" = "unwillingness-reason-I don’t often participate in conversations like these",
  "Might struggle to explain views or seem uninformed" = "unwillingness-reason-I might struggle to explain my views to the others and/or come across as uninformed",
  "Others might be unproductive or disrespectful" = "unwillingness-reason-I would not trust the others to keep the conversation productive and respectful (e.g., keeping emotions in check, refraining from making hostile remarks, not dominating the conversation)",
  "Views wouldn't change" = "unwillingness-reason-It would be a waste of time because my views wouldn't change",
  "Don't want to listen to others" = "unwillingness-reason-I would not want to listen to the others discuss their views",
  "Might be criticized" = "unwillingness-reason-Someone could criticize or show disapproval of my views",
  "Might experience negative feelings about the world" = "unwillingness-reason-The conversation could negatively affect how I feel about the world during and/or afterward",
  "Conversation might be awkward or tense" = "unwillingness-reason-The conversation could become awkward or tense",
  "Might feel embarrassed" = "unwillingness-reason-I could feel humiliated or embarrassed in the conversation",
  "Might offend someone" = "unwillingness-reason-Someone might take offense to what I say",
  "Might struggle to remain productive and respectful" = "unwillingness-reason-I might struggle to remain productive and respectful",
  "Others won't make effort to listen" = "unwillingness-reason-The others might not make a full effort to hear and understand me",
  "Identity might be challenged" = "unwillingness-reason-Someone could challenge ideas or beliefs that play an important role in making me who I am",
  "Might feel invalidated" = "unwillingness-reason-I might feel disempowered, unheard, or invalidated",
  "Might incur social repercussions" = "unwillingness-reason-I could incur social repercussions (e.g., being excluded by others in the future, negative changes to my reputation, feeling ostracized)",
  "Opinion is unpopular in general" = "unwillingness-reason-My opinion is unpopular in general",
  "Topic is too complex" = "unwillingness-reason-I would feel like the topic is too complex or requires more knowledge than I have",
  "No first-hand experience or relevance" = "unwillingness-reason-I don't have first-hand experience with the topic and/or it isn't relevant to me",
  "Non-social repercussions (e.g., work)" = "unwillingness-reason-I could experience significant repercussions at work or in other spaces beyond my personal social network",
  "Too mentally draining" = "unwillingness-reason-It wouldn't be worth the emotional and/or mental energy",
  "Others' views won't change" = "unwillingness-reason-It would be a waste of time because the others' views wouldn't change",
  "Might damage relationships" = "unwillingness-reason-I would not want to hurt my relationship(s) with others or upset them"
) %>%
  mutate(across(everything(), ~ replace_na(.x, 1))) # making the concerns that participants weren't even worried about into a 1 so that they can be represented in the likert graph as "not at all"


wide_unwill_reasons_df_2 <- wide_unwill_reasons_df_1 %>%
  select(-prolific_subject_id) %>%
  mutate(across(everything(), ~ factor(.x, levels = 1:7,
     labels = c("Not at all", "Slightly", "Somewhat",
                "Moderately", "Quite a bit", "Very much", "Extremely"),
     ordered = TRUE))) 

unwillingness_reason_likert <- likert(as.data.frame(wide_unwill_reasons_df_2))

# for the future when comparing pre and post, might want to use this example line: df1_likert <- likert(items=df1[,3:4], grouping=df1[,2])

plot(
  unwillingness_reason_likert, 
  legend.position="right",
  colors = c("#63666A", "#9A9B9C", "#D0D0CE", "#FFF8EF", "#F2D4D7", "#B2182B", "#641A2B")) +
  labs(title = "What Concerns Deter Americans from Engaging in Difficult Conversations?",
       subtitle = "Figure includes all subjects ratings of all 25 items") +
  theme(
    plot.title = element_text(
      hjust = 0.5,      # center
      size = 10         # font size
    )
  ) 

```


```{r}
#| label: find avg rating of endorsed concerns
#| echo: false

avg_rating <- surveys_df %>%
  filter(phase == "follow-up-unwillingness-concerns-survey") %>%
    # Keep only the columns needed
    select(question, answer) %>%
    # Group by the question text
    summarize(mean = round(mean(as.numeric(answer), na.rm = TRUE), 1))
```

As seen in @fig-unwillingness-reason-likert-bar-plot, no single item was a concern for the overwhelming majority of the sample. Rather, most items were not concerns at all for the overwhelming majority. Part of this can be explained simply by  the fact that the average participant was only concerned with a small subset of items. On average, subjects endorsed `r round(sum_stats_concerns_endorsed$mean, 0)` concerns as impacting their willingness at all (Min = `r sum_stats_concerns_endorsed$min`, LQ = `r sum_stats_concerns_endorsed$lower_quartile`, UQ = `r sum_stats_concerns_endorsed$upper_quartile`, Max = `r sum_stats_concerns_endorsed$max`). Of the concerns impacting their willingness, subjects, on average, rated a given concern's impact at `r avg_rating$mean` (4 = "Moderately" and 5 = "Quite a bit") out of the 7-pt Likert scale. 
 
 
```{r}
#| label: fig-unwillingness-reason-only-endorsed-likert-bar-plot
#| echo: false
#| fig-cap: Likert bar plot displaying subjects' ratings of how much certain concerns contribute to their unwillingness to converse. The 25 potential concerns were derived from existing literature and pretested in a pilot experiment. On a scale from 1 to 7, where a rating of 1 meant participant's willingness is not at all impacted by the concern and a rating of 7 meant participant's willingness is strongly impacted by the concern, participants rated the potential concerns' effects on their willingness to converse. This plot only shows ratings of 2 and greater to display the distributions of concern severity when a concern does impact willingness.

wide_unwill_reasons_df_3 <- wide_unwill_reasons_df_2 %>%
 mutate(across(everything(), ~ na_if(.,"Not at all")))

names(wide_unwill_reasons_df_3) <- str_wrap(names(wide_unwill_reasons_df_3), width = 40)
  
unwillingness_reason_only_endorsed_likert <- likert(as.data.frame(wide_unwill_reasons_df_3))

# for the future when comparing pre and post, might want to use this example line: df1_likert <- likert(items=df1[,3:4], grouping=df1[,2])

suppressWarnings(plot(
  unwillingness_reason_only_endorsed_likert, 
  legend.position="right",
# plot.percents = TRUE,
  colors = c("#63666A", "#9A9B9C", "#D0D0CE", "#FFF8EF", "#F2D4D7", "#B2182B", "#641A2B")) +
  labs(title = "How Severely do Americans' Concerns Deter Them from Engaging in Difficult Conversations?",
       subtitle = "Figure only includes subjects' ratings for items they endorsed as affecting their willingness in binary phase") +
  theme(
  plot.title = element_text(hjust = 0.5, size = 8),
  plot.subtitle = element_text(hjust = 0.5, size = 7),
    # reduce font size of y-axis to fit more items
    axis.text.y = element_text(size = 8, lineheight = 0.8),
    # add a larger left margin (t, r, b, l) so text doesn't cut off
    plot.margin = margin(10, 10, 10, 20)
    )
  )

```

@fig-unwillingness-reason-only-endorsed-likert-bar-plot demonstrates that when a participant endorses an item in the binary phase, meaning it is of any concerns to them at all, it more often than not plays at least a moderate role in shaping their willingness to converse. 

The top five most frequently endorsed concerns in the binary phase were: `r paste0(shQuote(top_5_concerns[1:5]), collapse=", ")`.

The top five concerns most severely impacting willingness were `r paste0(shQuote(top_5_severe[1:5]), collapse=", ")`. The full distribution of severity ratings can be observed in @fig-unwillingness-reason-likert-bar-plot.

Thus, while there are differences between the most endorsed and the most influential items, a common theme begins to surface: futility and cost of emotional/mental resources. 

#### Free response

To assess the comprehensiveness of the inventory, we analyzed the free-response data using Anthropic's Claude with manual review on 10% of the sample for validation. Claude detected select potentially unique concerns, but these were largely one-offs. The free response data suggest that our survey sufficiently captures the primary concerns driving Americans' reluctance to engage in conversations on controversial topics. 

### Profiling 

#### Factor Analysis on Concerns

We examined the initial binary endorsement ratings of our concern items where each participant simply indicated whether an item concerned them at all, as well as the Likert-scale ratings of how much each of the endorsed items impact the participant's willingness.

```{r}
#| label: tbl-free-response
#| include: false
#| fig-cap: Table of example free responses in which participants expressed avoidance of conversations because they felt that the topic was too complex or that they did not know enough about the topic to engage in a conversation about it.


# fr_knowledge <- free_response_df %>%
#   filter(prolific_subject_id == "67ac21373a76570ef0a5c0d3" | prolific_subject_id == "66b979589d208d3ec691d89b" | prolific_subject_id == "5d3695969749cc00165c222f") %>%
#   select("Free response", "Focus topic")

# kable(fr_knowledge, "html") %>%
#   kable_styling() %>%
#   column_spec(2, width = "5em")

```

```{r}
#| label: Wrangle data for factor analysis
#| echo: false

# join short labels back to the main data
surveys_df <- surveys_df %>%
  full_join(question_key_for_fa, by = "question")

concerns_labeled_wide <- surveys_df %>%
  filter(phase == "follow-up-unwillingness-concerns-survey") %>%
  select(prolific_subject_id, short_label, answer) %>%
  pivot_wider(names_from = short_label, values_from = answer) %>%
  mutate(across(starts_with("Q0"), ~ as.numeric(as.character(.)))) %>%
  ungroup() 

concerns_labeled_wide_na_replaced <- surveys_df %>%
  filter(phase == "follow-up-unwillingness-concerns-survey") %>%
  select(prolific_subject_id, short_label, answer) %>%
  pivot_wider(
    names_from = short_label, 
    values_from = answer,
    values_fill = "1"  
  ) %>%
  mutate(across(starts_with("Q0"), ~ as.numeric(as.character(.))))

concerns_wide_na_replaced <- concerns_labeled_wide_na_replaced %>%
  select(starts_with("Q0")) %>% 
  mutate(across(starts_with("Q0"), ~ as.numeric(as.character(.))))

concerns_wide <- concerns_labeled_wide %>%
  select(-prolific_subject_id)

concerns_matrix_numeric <- apply(concerns_wide_na_replaced, 2, function(x) as.numeric(as.character(x))) # r won't convert to numeric for some reason

row_na_counts <- rowSums(is.na(concerns_labeled_wide))
total_na <- sum(row_na_counts)
avg_na_per_row <- mean(row_na_counts)

```

```{r}
#| label: fig-correlation-matrix
#| echo: false
#| fig-cap: Correlation matrix of reasons for unwillingness

cor_matrix <- cor(concerns_wide_na_replaced, use = "pairwise.complete.obs")

library(corrplot)
corrplot <- corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.7)

# the items all seem to show low to moderate correlations. pretty solid

```

@fig-correlation-matrix indicates a mix of zero to moderate positive correlation strengths between various willingness concern items. A few pairings had negative correlations, but all were very minimal. This range is consistent with the expectation that many of these concerns would overlap in some ways but that no two items represent the same concern. 

```{r}
#| label: Check KMO, Bartlett's test, and Rit values to see if data are well-suited for factor analysis 
#| include: false 

# KMO and Bartlett’s test
kmo <- KMO(concerns_wide_na_replaced)
cortest_bartlett <- cortest.bartlett(cor(concerns_wide_na_replaced), n = nrow(concerns_wide_na_replaced))

#overall msa is above 0.8 so our data are well-suited for factor analysis!  

#cortest_bartlett shows that the p-value is VERY small, so correlations are statistically significant 

#check internal consistency
rit.vals <- alpha(concerns_wide_na_replaced)

# all indicators suggest that the data are well-suited for factor analysis 

```

```{r}
#| label: Factor analysis on 25 items 
#| include: false

# determine number of factors with parallel analysis. this will output a suggested number of factors and a scree plot 
fa_parallel_results <- fa.parallel(concerns_wide_na_replaced, fa = "fa")

# the results of parallel analysis suggest 4 factors 

# run factor analysis with 4 factors 
## we use oblimin because the factors are probably correlated 
fa_result <- fa(
    concerns_wide_na_replaced,
    nfactors = 4,
    rotate = "oblimin"
    ) 

#get explained variance for report 
total_var_explained_percent <- round(sum(fa_result$Vaccounted["Proportion Var", ]) * 100, 1)

#get factor correlations 
factor_corrs <- fa_result$r 

# extract min and max off-diagonal correlations
cor_range <- range(factor_corrs[lower.tri(factor_corrs)])
cor_min <- round(cor_range[1], 2)
cor_max <- round(cor_range[2], 2)
```

```{r}
#| label: tbl-factor-loadings-visual
#| echo: false
#| tbl-cap: Heat-mapped-table visualizing factor loadings in exploratory factor analysis. I used parallel analysis to choose four factors. 

# visualize factor loadings
#prepping dfs
loadings_full_df <- as.data.frame(fa_result$loadings[1:25, ]) %>%
  dplyr::rename("1" = "MR1",
         "2" = "MR2",
         "3" = "MR3",
         "4" = "MR4")

loadings_full_df$short_label <- rownames(loadings_full_df)

#pivot for readability
loadings_full_long <- loadings_full_df %>% 
  pivot_longer(-short_label, names_to = "factor", values_to = "loading")

# make digestible df with only the top loading items so we can see which concerns are in each factor
loadings_max_df <- loadings_full_long %>%
  group_by(short_label) %>%
  slice_max(order_by = abs(loading), n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  merge(question_key_for_fa, by = "short_label") %>%
  arrange(factor)

#make plot for visualization 
loadings_full_long <- loadings_full_long %>%
  merge(question_key_for_fa, by = "short_label")

factor_loadings_plot <- ggplot(loadings_full_long, aes(x = reorder(short_concern_label, loading), y = loading, fill = factor)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(
    title = "Factor Loadings",
    x = "Concern",
    y = "Loading",
    fill = "Factor"
  ) +
  theme_minimal()

#pull items for printing later
factor_1_items <- loadings_max_df %>%
  filter(factor == "1") %>%
 pull(short_concern_label)  

factor_2_items <- loadings_max_df %>%
  filter(factor == "2") %>%
 pull(short_concern_label)  

factor_3_items <- loadings_max_df %>%
  filter(factor == "3") %>%
 pull(short_concern_label)  

factor_4_items <- loadings_max_df %>%
  filter(factor == "4") %>%
 pull(short_concern_label)  

#make item order for printing
item_order <- loadings_max_df %>%
  arrange(factor, desc(loading)) %>%
  pull(short_concern_label)

# print both ways
#factor_loadings_plot

# use gt lib for pretty printing
loadings_full_long %>% 
  select(factor, loading, short_concern_label) %>%
  pivot_wider(names_from = factor, values_from = loading) %>%
  arrange(match(short_concern_label, item_order)) %>%
  gt() %>%
  cols_label(short_concern_label = "Concern") %>%
  fmt_number(decimals = 2) %>%
  data_color(
    columns = where(is.numeric),
    method = "numeric",
    palette = c("blue", "white", "red"),
    domain= c(-1, 1)
  )

```

```{r}
#| label: characterizing factor scores 
#| echo: false
#| warning: false
#| message: false

# make factor scores into a df and match up with subject id's from the df i used to run factor analysis
factor_scores_df <- as.data.frame(fa_result$scores) %>%
  mutate(prolific_subject_id = concerns_labeled_wide_na_replaced$prolific_subject_id) %>%
  rename(
    "1" = "MR1",
         "2" = "MR2",
         "3" = "MR3",
         "4" = "MR4")

# now summarize factor scores across sample
standardized_summary_factor_scores_df <- factor_scores_df %>%
  summarise(across(all_of(c("1", "2", "3", "4")),
                   list(mean = mean, sd = sd),
                   na.rm = TRUE))

super_summary_factor_scores_df <- factor_scores_df %>%
  pivot_longer(all_of(c("1", "2", "3", "4")),
               names_to = "factor",
               values_to = "score") %>%
  group_by(prolific_subject_id) %>%
  slice_max(score, n = 1) %>%
  ungroup() %>%
  count(factor) %>%
  mutate(pct = n / n_distinct(factor_scores_df$prolific_subject_id))
```

```{r}
#| label: characterize sample esndorsement and ratings of factors 
#| echo: false
#| warning: false
#| message: false

## data prep
#merge with big df 
surveys_df <- surveys_df %>%
  left_join(loadings_max_df %>% select(question, factor), by = "question")

# separate out endorsement and severity from the answer column
endorsements <- surveys_df %>%
  filter(phase == "willingness-concerns-first-survey") %>%
  select(prolific_subject_id, question, endorsed = answer)

severity <- surveys_df %>%
  filter(phase == "follow-up-unwillingness-concerns-survey") %>%
  select(prolific_subject_id, question, severity = answer)


# now count number of items in each factor 
factor_endorsement_summary <- endorsements %>%
  left_join(loadings_max_df %>% select(question, factor), by = "question") %>%
  filter(endorsed == 1) %>%  # 1 means yes, 2 means no
  group_by(prolific_subject_id, factor) %>%
  summarise(n_endorsed = n()) %>%
  ungroup() %>%
  group_by(factor) %>%
  summarise(
    pct_subjects_endorsing_any = n_distinct(prolific_subject_id) / n_distinct(endorsements$prolific_subject_id),
    mean_items_endorsed = mean(n_endorsed)
  ) %>%
  mutate(n_items_in_factor = case_when(
    factor == "1" ~ length(factor_1_items),
    factor == "2" ~ length(factor_2_items),
    factor == "3" ~ length(factor_3_items),
    factor == "4" ~ length(factor_4_items),
    .default = NA
  ))

```

```{r}
#| label: fig-fa-heat-map-df
#| echo: false
#| tbl-cap: Table showing heat map of factor analysis on concern items.

fa_heat_map <- ggplot(loadings_full_long, aes(
  x = factor, 
  y = factor(short_concern_label, levels = item_order), 
  fill = loading)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = 0, limit = c(-1, 1), name = "Loading"
  ) +
  labs(
    title = "Heat Map of Factor Loadings",
    x = "Factor",
    y = "Concern"
  ) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10))

print(fa_heat_map)

```

We conducted an exploratory factor analysis (EFA) on the concern items using oblimin rotation to identify latent factors that could reflect correlated concerns. Parallel analysis suggested a four-factor solution, which accounted for `r total_var_explained_percent`% of the total variance. The factors were moderately correlated ($r$ = `r cor_min`–`r cor_max`) as shown in @tbl-factor-loadings-visual and @fig-fa-heat-map-df (which accomplish the same thing; I included @tbl-factor-loadings-visual for its descriptiveness and @fig-fa-heat-map-df for its visual ease). Many cross-loadings were present, consistent with the expectation of conceptual overlap among factors. We retained all items to allow full examination of the emerging structure.

If these factors reveal true overarching themes in the concern items, we interpret them as:

Social & Reputation Consequences (Factor 1): This factor reflects fears of external, social consequences. The items in this factor are: `r paste0(shQuote(factor_1_items), collapse=", ")`.

Perceived Futility & Emotional Cost (Factor 2): This factor reflects a sense that the conversation is not worth engaging in because it will not amount to anything or it will be too emotionally or mentally costly. The items in this factor are: `r paste0(shQuote(factor_2_items), collapse=", ")`.

Unpreparedness for Complex Topic (Factor 3): This factor captures doubts of knowledge about topic and ability to convey one's ideas. The items in this factor are: `r paste0(shQuote(factor_3_items), collapse=", ")`.

Personal Vulnerability & Identity Threat (Factor 4): This factor consists of items focusing on internal psychological distress and threats to the participant's self-concept or core beliefs. The items in this factor are: `r paste0(shQuote(factor_4_items), collapse=", ")`.

About `r round(100*(factor_endorsement_summary$pct_subjects_endorsing_any[factor_endorsement_summary$factor == "1"]), 0)` percent of subjects endorsed at least one concern in factor 1 (Social & Reputation Consequences). On average across the entire sample, subjects endorsed `r round(factor_endorsement_summary$mean_items_endorsed[factor_endorsement_summary$factor == "1"], 1)` of the `r factor_endorsement_summary$n_items_in_factor[factor_endorsement_summary$factor == "1"]` items in factor 1. Similarly, `r round(100*(factor_endorsement_summary$pct_subjects_endorsing_any[factor_endorsement_summary$factor == "2"]), 0)` percent of subjects endorsed at least one concern in factor 2 (Perceived Futility & Emotional Cost). On average across the entire sample, subjects endorsed `r round(factor_endorsement_summary$mean_items_endorsed[factor_endorsement_summary$factor == "2"], 1)` of the `r factor_endorsement_summary$n_items_in_factor[factor_endorsement_summary$factor == "2"]` items in factor 2. `r round(100*(factor_endorsement_summary$pct_subjects_endorsing_any[factor_endorsement_summary$factor == "3"]), 0)` percent of subjects endorsed at least one concern in factor 3 (Unpreparedness for Complex Topic). On average across the entire sample, subjects endorsed `r round(factor_endorsement_summary$mean_items_endorsed[factor_endorsement_summary$factor == "3"], 1)` of the `r factor_endorsement_summary$n_items_in_factor[factor_endorsement_summary$factor == "3"]` items in factor 3. Finally, `r round(100*(factor_endorsement_summary$pct_subjects_endorsing_any[factor_endorsement_summary$factor == "4"]), 0)`  percent of subjects endorsed at least one concern in factor 4 (Personal Vulnerability & Identity Threat). On average across the entire sample, subjects endorsed `r round(factor_endorsement_summary$mean_items_endorsed[factor_endorsement_summary$factor == "4"], 1)` of the `r factor_endorsement_summary$n_items_in_factor[factor_endorsement_summary$factor == "4"]` items in factor 4.

Factor scores were computed using the regression method. Factor 1 was the dominant concern domain for `r round(100*(super_summary_factor_scores_df$pct[super_summary_factor_scores_df$factor == "1"]), 0)` percent of participants. Factor 2 was the dominant concern domain for `r round(100*(super_summary_factor_scores_df$pct[super_summary_factor_scores_df$factor == "2"]), 0)` percent of participants. Factor 3 was the dominant concern domain for `r round(100*(super_summary_factor_scores_df$pct[super_summary_factor_scores_df$factor == "3"]), 0)` percent of participants. Factor 4 was the dominant concern domain for `r round(100*(super_summary_factor_scores_df$pct[super_summary_factor_scores_df$factor == "4"]), 0)` percent of participants.

#### PCA & Hierarchical Cluster Analysis

```{r}
#| label: PCA for profiling participants
#| include: false

# run pca on participants 
pca_result <- prcomp(concerns_wide_na_replaced, scale. = TRUE)
summary(pca_result)

#screeplot helps search for an elbow that indicates where the number of principal components should be cut off
screeplot(pca_result, type = "lines")

#biplot is hard to gain info from with this many observations and vectors 
biplot(pca_result, scale = 0)

## look at proportions of variance explained by each principal component 
pve <- summary(pca_result)[["importance"]]["Proportion of Variance",]
#plot
plot(1:length(pve), pve, ylim = c(0,1), xlab = "PC", 
     ylab = "Prop Variance Explained", type = "b", pch = 20)

## i find cumulative more helpful:
cumpve <- summary(pca_result)[["importance"]]["Cumulative Proportion",]
cumpve
#plot
plot(1:length(cumpve), cumpve, ylim = c(0,1), xlab = "PC", 
     ylab = "Cum Prop Variance Explained", type = "b", pch = 20)

# add pca_results to the og concerns df
concerns_labeled_wide_na_replaced <- data.frame(concerns_labeled_wide_na_replaced, pca_result$x)

```

```{r}
#| label: fig-pca-scree-plot
#| fig-cap: Scree plot indicating some structure in the principal components--- meaningful drop from PC1 to PC2, and more notable elbow from PC4 to PC5. Then all components are relatively negligible contributors. Suggests we should use 4 PCs.
#| echo: false 

screeplot(pca_result, type = "lines")
```

```{r}
#| label: look at pca loadings
#| echo: false

# grab four pc loadings 
four_pca_loadings <- as.data.frame(round(pca_result$rotation[, 1:4], 2)) %>%
  arrange(desc(PC2)) %>%
  mutate(PC1 = as.numeric(PC1),
         PC2 = as.numeric(PC2),
         PC3 = as.numeric(PC3),
         PC4 = as.numeric(PC4))

#switch out labels for the shortened text labels 
rownames(four_pca_loadings) <- question_key_for_fa$short_concern_label[match(rownames(four_pca_loadings), question_key_for_fa$short_label)]

# highlight the extreme loadings so i can pay closer attention 
extreme_pc1_loadings <- four_pca_loadings %>%
#  filter(abs(PC1) > 0.27) %>%
  arrange(desc(PC1)) %>%
  select(PC1)

extreme_pc2_loadings <- four_pca_loadings %>%
  filter(abs(PC2) > 0.27) %>%
  select(PC2)

extreme_pc3_loadings <- four_pca_loadings %>%
  filter(abs(PC3) > 0.27) %>%
  select(PC3)

extreme_pc4_loadings <- four_pca_loadings %>%
  filter(abs(PC4) > 0.27) %>%
  select(PC4)
```

We conducted a principal component analysis (PCA) on participants' concern rating profiles to examine the dimensionality of their responses. The scree plot (@fig-pca-scree-plot) indicated that four principal components captured the majority of explainable variance, and we focus our interpretation on these.

The first principal component, accounting for `r round(100*(pve[1]), 0)`% of variance, showed uniformly positive loadings across all items (range of 0.10–0.27). This pattern is consistent with a general response factor: participants who score high on PC1 rate all concerns highly, reflecting an overall tendency toward greater reluctance rather than concern about any particular aspect of conversation. The second, third, and fourth principal components all had at least four items that loaded onto the components with magnitudes of 0.29 and greater; the greatest loading on the first principal component is 0.27, with most items loading in the tenths. 

```{r}
#| label: tbl-pca-two-loadings 
#| echo: false

kable(extreme_pc2_loadings, caption = "Concern items with extreme PC2 loadings")

```

The second principal component accounted for `r round(100*(pve[2]), 0)`% of variance. Its highest-loading items reflect concerns about the futility of engagement and the anticipated emotional or mental costs outweighing potential benefits. This pattern is consistent with the  sentiments observed in @fig-unwillingness-reason-likert-bar-plot and @fig-unwillingness-reason-only-endorsed-likert-bar-plot and reflects factor 2 (Perceived Futility & Emotional Cost) from the factor analysis. 

```{r}
#| label: tbl-pca-three-loadings 
#| echo: false

kable(extreme_pc3_loadings, caption = "Concern items with extreme PC3 loadings")
```

The third principal component accounted for `r round(100*pve[3], 0)`% of variance and contrasted two distinct sources of reluctance. Positively loading items comprise factor 3 (Unpreparedness for Complex Topic) from factor analysis, reflecting a sense of being ill-equipped to engage due to topic complexity, limited first-hand experience, or difficulty articulating one's views. Negatively loading items reflected concern about holding an unpopular opinion and the social consequences that might follow from expressing it; both of these items factored into factor 1 (Social & Reputation Consequences) during factor analysis.

```{r}
#| label: tbl-pca-four-loadings 
#| echo: false

kable(extreme_pc4_loadings, caption = "Concern items with extreme PC4 loadings")
```

The fourth principal component accounted for `r round(100*pve[4], 0)`% of variance and captured a similar but distinct contrast. Negatively loading items reflected reluctance stemming from unfamiliarity with the topic and with navigating disagreement; these items all loaded into different factors during factor analysis. Positively loading items reflected avoidance motivated by anticipated negative affect, both self-directed and directed toward others; these items both fell into factor 4 (Personal Vulnerability & Identity Threat).

```{r}
#| label: Hierarchical cluster analysis
#| include: false

#calculate distances
concern_dists <- concerns_wide_na_replaced %>%
  dist()

#perform the clustering with average and complete linkage 
hclustmodaverage <- concern_dists %>% hclust(method = "average")
hclustmodcomplete <- concern_dists %>% hclust(method = "complete")

#plot full dendrograms
plot(hclustmodaverage)
plot(hclustmodcomplete)

#moving forward with complete linkage tree because it is more distributed than average linkage

plot(hclustmodcomplete) %>%
abline(h = 24, col = "blue", lty = 2)

# cutting at 3 because it is as even as i can get the clusters to be
completecut4 <- cutree(hclustmodcomplete, k = 4)
table(completecut4)

# add cluster id back to the df 
concerns_with_cluster_df <- concerns_labeled_wide_na_replaced %>% mutate(clusterID = completecut4) 

#get percentages to see the distribution amongst the clusters 
hierarchical_cluster_percentages <- concerns_with_cluster_df %>%
  group_by(clusterID) %>%                       # group by cluster
  summarise(n = n()) %>%  # count unique subjects
  mutate(percent = round(n / sum(n) * 100, 1))       # calculate percentage

# add to surveys_df
hierarchical_cluster_info <- concerns_with_cluster_df %>%
  select(prolific_subject_id, clusterID)

surveys_df <- surveys_df %>%
  left_join(hierarchical_cluster_info, by = "prolific_subject_id")

```

```{r}
#| label: fig-dendrogram-hierarchical-clustering
#| echo: false
#| fig-cap: A dendrogram cut at four clusters shows the results of hierarchical cluster analysis using complete linkage and Euclidean distance to look for common concern profiles amongst participants. 

# print the dendrogram with the cut that i selected in previous code chunk 
plot(hclustmodcomplete) %>%
abline(h = 22, col = "blue", lty = 2)

```

I also conducted hierarchical clustering (euclidean distance, complete linkage), on participants' ratings of concerns to determine whether there are trends in participants ratings that could form general participant concern profiles (@fig-dendrogram-hierarchical-clustering). I chose to use four clusters primarily to avoid extremely small clusters (see the cluster to the far right in @fig-dendrogram-hierarchical-clustering). As seen in @fig-dendrogram-hierarchical-clustering, many participants fell into one cluster; this pattern was more extreme in the case of single linkage, which motivated my choice of complete linkage. 

The trend of one cluster or factor comprising a large proportion of participants continually recurred throughout my analyses, and was one of the major motivations for my use of so many profiling methods.

```{r}
#| label: fig-heat-map-hierarchical-clustering-on-items
#| echo: false
#| fig-cap: Heat map of each cluster's average rating of each concern item. Clusters were formed through hierarchical cluster analysis using complete linkage and Euclidean distance.

#make df for heatmap 
hierarchical_for_heatmap_df <- concerns_with_cluster_df %>%
  group_by(clusterID) %>%
  summarise(across(starts_with("Q"), mean, na.rm = TRUE)) %>%
  pivot_longer(cols = c(Q01,Q02,Q03,Q04,Q05,Q06,Q07,Q08,Q09,Q010,Q011,Q012,Q013,Q014,Q015,Q016,Q017,Q018,Q019,Q020,Q021,Q022,Q023,Q024,Q025), names_to = "short_label", values_to = "mean_rating") %>%
  merge(question_key_for_fa, by = "short_label")

#get order to sort clusters in heatmap 
heat_map_concern_order <- hierarchical_for_heatmap_df %>%
  filter(clusterID == 1) %>%
  arrange(mean_rating) %>%
  pull(short_concern_label)

#make heat map to see what the cluster profiles actually look like  
hierarchical_clustering_heatmap <- ggplot(
  hierarchical_for_heatmap_df, 
    aes(
      x = factor(clusterID), 
      y = factor(short_concern_label, 
      levels = heat_map_concern_order), 
      fill = mean_rating)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "#2f2fff",     # blue
    mid = "#f2f2f2",     # light center
    high = "#ff2a1a",    # red
    midpoint = 3,
    limits = c(1, 7),   # matches your example scale
    breaks = c(1, 2, 3, 4, 5, 6, 7)
  ) + 
  labs(x = "Participant Profile", 
       y = "Concern", 
       fill = "Mean Rating", 
       title = "Four Distinct Profiles of Conversational Reluctance") +
  theme(axis.text.y = element_text(size = 8))  # no more angle needed

hierarchical_clustering_heatmap

# look at pc scores by cluster
 pc_scores <- as.data.frame(pca_result$x[, 1:4])
 pc_scores$clusterID <- concerns_with_cluster_df$clusterID
 
 # this is printable
pc_scores_by_cluster <- pc_scores %>%
   group_by(clusterID) %>%
   summarise(across(everything(), mean))

```

The majority of participants, `r hierarchical_cluster_percentages$percent[hierarchical_cluster_percentages$clusterID == 1]`%, fell into cluster 1, averaging relatively low concern across most items (@fig-heat-map-hierarchical-clustering-on-items). This suggests that most participants in the sample self-reported as endorsing few concerns and not showing meaningful reluctance towards any particular aspect of conversing. This somewhat aligns with PCA's first principal component, which captures the notion that one primary profiling dimension is participants' severity of ratings overall. 

A further `r hierarchical_cluster_percentages$percent[hierarchical_cluster_percentages$clusterID == 2]`% of participants fell into cluster 2. These participants overlapped with much of cluster 1's lack of concern for many items; however, there were some differences in what did not concern cluster 2 compared to cluster 1; additionally, cluster 2 showed greater concern for many items that cluster 1 rated as only slightly concerning. Cluster 2 showed a relative elevation on items related to views changing (both their own and others'), emotional and mental costs, conversation awkwardness, and conversation futility.

Cluster 3 participants, `r hierarchical_cluster_percentages$percent[hierarchical_cluster_percentages$clusterID == 3]`% of the sample, reported moderate-to-high concern broadly with particular elevation on items reflecting a sense of being ill-equipped to engage. This included struggling to explain their views, feeling invalidated, and finding the topic too complex. Their reluctance appears driven primarily by fear of others being disrespectful and not giving the participant the psychological safety and respect needed to imperfectly navigate the topic, as opposed to concerns like listening to others.  

Cluster 4 participants reported the highest concern overall, with especially strong ratings on items reflecting anticipated futility: that the conversation would be too mentally draining, that they and others would struggle to remain respectful and productive, and that neither their own nor others' views would change as a result. Unlike cluster 3's suspicion that the others will not give them the space to navigate the topic and be heard, cluster 4's reluctance seems bidirectional and focused on productivity. Just `r hierarchical_cluster_percentages$percent[hierarchical_cluster_percentages$clusterID == 4]`% of participants fell into cluster 4.

Taken together, the four clusters may suggest a sort of floor effect on severity ratings that compresses most participants into the low-concern region and prevents finer differentiation by concern types. Overall level of reluctance seems to be the driving factor that captures a participant in cluster 1 and potentially even cluster 2, the largest clusters. The two smaller clusters then have subsets of concerns surface: cluster 3 is concerned about competency in the conversation and others not listening nor allowing for imperfection, whereas cluster 4 is skeptical of both parties' abilities to maintain a productive atmosphere and reach a useful conclusion.

These findings should be interpreted in light of the methodological consideration that although participants rated all 25 concern items, on average, only `r round(sum_stats_concerns_endorsed$mean, 0)` concerns actually affected a given participant's willingness to engage. I chose to impute the non-endorsed concerns as 1s so that I could conduct analyses like factor analysis and PCA; however, this may have deflated mean ratings across the board, potentially pushing participants toward the lower-concern clusters and even masking undetected concern profiles. The low ratings in Clusters 1 and 2 may therefore reflect not an absence of reluctance per se, but rather that most participants' reluctance is relatively selective. This, as well as the small sizes of clusters 3 and 4, could possibly be explored further through bootstrap sampling or probabilistic PCA (need to learn more about probabilistic PCA).
 
```{r}
#| label: fig-pca-cluster-scatterplot
#| echo: false
#| fig-cap: Plot displaying participant clusters by their scores on the first and second principal components found through principal component analysis (PCA). We used hierarchical cluster analysis with Euclidean distance and complete linkage to create four clusters of participants based on their concern profile. The first principal component is likely a general factor, where participants score higher on this component if they generally rated the 25 concern items higher. The second principal component appears to differentiate between high scorers who are concerned about the futility of engagement and the anticipated emotional or mental costs outweighing potential benefits, and others who scored moderately.

pc_scores_df <- as.data.frame(pca_result$x[, 1:2])
pc_scores_df$clusterID <- factor(concerns_with_cluster_df$clusterID)

ggplot(pc_scores_df, aes(x = PC1, y = PC2, color = clusterID)) +
  geom_point(alpha = 0.6, size = 2) +
  scale_color_brewer(palette = "Set2", name = "Cluster") +
  labs(
    title = "Participant Concern Profiles in PC Space",
    x = "PC1 (Overall Concern Level)",
    y = "PC2 (Futility vs. Social Concerns)"
  ) +
  theme_minimal()

```

@fig-pca-cluster-scatterplot demonstrates that the four clusters occupy meaningfully distinct regions of the PCA space, with PC1 (overall concern level) serving as the primary axis of separation. Cluster 1 sits at the low end of PC1, consistent with its broadly low concern profile, while cluster 4 sits at the high end, consistent with its broadly elevated concern profile. Cluster 2 separates from cluster 1 primarily along PC2 rather than PC1, indicating that these participants might not be substantially more concerned overall but are elevated on futility- and emotional-cost-related items. This corroborates the cluster-mean pattern in @fig-heat-map-hierarchical-clustering-on-items. Cluster 3 occupies greater PC1 values without strong PC2 elevation, consistent with its profile being defined by a different component and relatively greater overall concern than clusters 1 and 2. Taken together, this projection supports the interpretation that overall reluctance level (PC1) is the dominant axis distinguishing participants, with futility-emotional cost concerns (PC2) acting as a secondary differentiator that primarily distinguishes Cluster 2 from Cluster 1.

```{r}
#| label: extra text 
#| include: false

#Cluster 1's average rating of a given concern was `r cluster_means$overall_mean[cluster_means$cluster == 1]`. On average, Cluster 1 rated the following concerns the highest: `r paste(shQuote(cluster1_top), collapse = "; ")`.

#Cluster 2's average rating of a given concern was `r cluster_means$overall_mean[cluster_means$cluster == 2]`. On average, cluster 2's top concerns were `r paste(shQuote(cluster2_top), collapse = "; ")`.

#Cluster 3's average rating of a given concern was `r cluster_means$overall_mean[cluster_means$cluster == 3]`. On average, cluster 3's top concerns were `r paste(shQuote(cluster3_top), collapse = "; ")`.

```

```{r}
#| label: fig-hierarchical-fig-clusters-avg-willingness   
#| echo: false 
#| fig-cap: Box plot of average willingness to converse about topic by cluster per hierarchical cluster analysis.

avg_willingness_cluster_hierarchical_df <- avg_willingness_df %>%
  merge(concerns_with_cluster_df, by = "prolific_subject_id")

cluster_willingness_boxplot <-
ggplot(avg_willingness_cluster_hierarchical_df, aes(x = factor(clusterID), y = avg_willingness, fill = factor(clusterID))) +
  geom_boxplot() +
  labs(x = "Participant Concern Profile (4 Clusters)", y = "Average Willingness to Converse (0-100)", title = "Willingness to Engage in Conversation by Concern Profile") +
  theme_minimal() +
  geom_text(
    data = hierarchical_cluster_percentages,
    aes(
      y = Inf,
      x = as.factor(clusterID),
      label = paste0("n = ", n)
    ),
    inherit.aes = FALSE,
    hjust = 0.5,
    vjust = 1,
    size = 3
  ) +
  scale_fill_manual(values = c("1" = "#951829", "2" = "#951829", "3" = "#951829", "4" = "#951829")) +
    guides(fill = "none")

cluster_willingness_boxplot

```

@fig-hierarchical-fig-clusters-avg-willingness generally corroborates the average valences of clusters as seen in @fig-heat-map-hierarchical-clustering-on-items: Cluster 1's lack of concern severity corresponds to a higher willingness to converse and cluster 4's great concern severity corresponds to a lower willingness to converse. Cluster 2's wide spread and lesser median compared to cluster 3 may be indicative of cluster 2's role as somewhat of a bridge between the lack of concern from cluster 1 and the strong but specialized sets of concerns in clusters 3 and 4. 

#### Latent Profile Analysis

I intend to conduct latent profile analysis after graduation.

### Open-mindedness & Willingness Relationship

```{r}
#| label: fig-openmindedness-likert-bar-plot
#| echo: false
#| warning: false
#| message: false 
#| fig-cap: Likert bar plot displaying subjects' ratings of five questions about their open-mindedness. The five questions were derived from @elnakouri2024 and modified for application to this experiment. On a scale from 1 to 7, where a rating of 1 meant subject strongly disagreed and a rating of 7 meant strongly agree, subjects rated how true the five statements were for them. Three of the statements were negatively worded in the survey but have been worded positively and reverse-coded in this plot for clarity.

# only works though if all "1" to "7" ratings are in the df

wide_for_likert_openmindedness_df_with_sub_labels <- openmindedness_df %>%
  select(question, answer, prolific_subject_id) %>%
  pivot_wider(names_from = question, values_from = answer) %>%
  ungroup() %>%
  rename("Has patience for opposing arguments" = "openmindedness...have little patience for arguments that I disagree with @R@", 
"Does not avoid opposing messages" = "openmindedness...avoid messages that I disagree with @R@", 
"Believes it is important to pay attention to all politics ideas" = "openmindedness...believe it is a waste of time to pay attention to certain political ideas about $topic @R@",
"Is open to other political viewpoints" = "openmindedness...am open to considering other political viewpoints about $topic",
"Considers as many different opinions as possible" = "openmindedness...consider as many different opinions as possible about $topic") 

wide_for_likert_openmindedness_df <- wide_for_likert_openmindedness_df_with_sub_labels %>%
  select(-prolific_subject_id) %>%
  mutate(across(everything(), ~ factor(.x, levels = 1:7,
     labels = c("Strongly Disagree", "Moderately Disagree", "Somewhat Disagree",
                "Neutral", "Somewhat Agree", "Moderately Agree", "Strongly Agree"),
     ordered = TRUE)))

### create a vector of the all 1-7 values. defining again here in case stuff gets moved around 

all_likert_values <- c("Strongly Disagree", "Moderately Disagree", "Somewhat Disagree", "Neutral", "Somewhat Agree", "Moderately Agree", "Strongly Agree")
### get the column names from  existing df
openmindedness_column_names <- colnames(wide_for_likert_openmindedness_df)
### create 7 new rows
new_openmindedness_likert_rows <- lapply(all_likert_values, function(response) {
  as.list(setNames(rep(response, length(openmindedness_column_names)), openmindedness_column_names))
}) %>%
  bind_rows()
### add the new rows to the existing dataframe
wide_for_likert_openmindedness_df <- bind_rows(wide_for_likert_openmindedness_df, new_openmindedness_likert_rows)


# standardize all columns to use the same factor object
wide_for_likert_openmindedness_df <- wide_for_likert_openmindedness_df %>%
  mutate(across(everything(), ~ factor(.x, levels = all_likert_values, ordered = TRUE)))

colnames(wide_for_likert_openmindedness_df) <- 
  str_wrap(colnames(wide_for_likert_openmindedness_df), width = 20)


openmindedness_likert <- likert(as.data.frame(wide_for_likert_openmindedness_df))

# for the future when comparing pre and post, might want to use this example line: df1_likert <- likert(items=df1[,3:4], grouping=df1[,2]) 

suppressWarnings(plot(openmindedness_likert, , 
  legend.position="right",
  colors = c("#63666A", "#9A9B9C", "#D0D0CE", "#FFF8EF", "#F2D4D7", "#B2182B", "#641A2B")) +
  labs(title = "How Open-Minded Are Participants?") +
  theme(
    plot.title = element_text(
      hjust = 0.5,      # center
      size = 10         # font size
    )
  ) )
```


On average, subjects rated their open-mindedness at `r mean_openmindedness` out of 7 (Range of `r min_openmindedness`-`r max_openmindedness`), where 1 is less open-minded and 7 is more open-minded, as observed in @fig-openmindedness-likert-bar-plot. Given the prevalence of ideas related to emotional and mental cost, conversation futility, and lack of expertise on complex topics, as well as indications that many participants anticipate that others will not listen to the participants' views, it is notable here that "Does not avoid opposing messages" and "Believes it is important to pay attention to all political ideas" respectively have the first and second lowest open-mindedness ratings from participants. 


```{r}
#| label: fig-EDA-correlation-openminded-relationships 
#| include: false 
#| warning: false
#| message: false
#| fig-cap: Relationships between each willingness concern and average open-mindedness. Using a 7-pt Likert scale, participants rated the impact of each willingness concern on their overall willingness. Also using a 7-pt Likert scale, participants rated 5 statements about their open-mindedness, which we averaged to find each participant's average open-mindedness.  

library(gridExtra) # so we can print the plots in columns 

cor_openmindedness_mean_vs_concerns <- concerns_wide_na_replaced %>% map_dfr(~ tidy(cor.test(.x, openmindedness_summary_by_subject_df$mean_openmindedness)), .id = "concern")

concerns_and_avg_openmindedness_df <- openmindedness_summary_by_subject_df %>%
  select(mean_openmindedness, prolific_subject_id) %>%
  full_join(concerns_labeled_wide_na_replaced, by = "prolific_subject_id") %>%
  select(-"prolific_subject_id")

# see the average number of concerns that people indicated as affecting them 

# Create all the individual plots
p1.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q01, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p2.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q02, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p3.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q03, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p4.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q04, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p5.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q05, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p6.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q06, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p7.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q07, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p8.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q08, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p9.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q09, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p10.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q010, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p11.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q011, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p12.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q012, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p13.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q013, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p14.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q014, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p15.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q015, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p16.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q016, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p17.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q017, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p18.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q018, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p19.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q019, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p20.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q020, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p21.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q021, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p22.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q022, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p23.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q023, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p24.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q024, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p25.o <- ggplot(concerns_and_avg_openmindedness_df, aes(x = Q025, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0)


# arrange the plots in a grid
#suppressWarnings(grid.arrange(p1.o, p2.o, p3.o, p4.o, p5.o, p6.o, p7.o, p8.o, p9.o, p10.o, p11.o, p12.o, p13.o, p14.o, p15.o, p16.o, p17.o, p18.o, p19.o, p20.o, p21.o, p22.o, p23.o, p24.o, p25.o, ncol = 4))


### the results seem to pretty much show no relationship, but if you squint your eyes, there is vaguely a linear relationship in some of the variables. i'm going to stick with pearson because, to my knowledge, the data doesn't meet the standards for any non-parametric methods 

#print example plot since printing all of them is way too much
ggplot(concerns_and_avg_openmindedness_df, aes(x = Q011, y = mean_openmindedness)) + geom_point() + geom_jitter(width = 0.2, height = 0) + 
  labs(x = "Rating of concern item's affect on willingness", y = "Open-mindedness rating", title = "Open-mindedness vs. severity of concern's effect on willingness")

```

```{r}
#| label: Correlations open-mindedness average vs. willingness concerns (NA is not replaced)
#| include: false
#| tbl-cap: Correlation matrix of average open-mindedness vs. willingness concerns 

# look at correlation relationship between mean open-mindedness score and each willingness concern
library(broom)

cor_openmindedness_mean_vs_concerns <- concerns_wide_na_replaced %>% map_dfr(~ tidy(cor.test(.x, openmindedness_summary_by_subject_df$mean_openmindedness)), .id = "concern")

# view results
cor_openmindedness_mean_vs_concerns_simplified <- cor_openmindedness_mean_vs_concerns %>%
  select(concern, estimate, p.value) %>%
  arrange(p.value) %>% 
  full_join(question_key_for_fa, by = c("concern" = "short_label"))
```

```{r}
#| label: Correlations open-mindedness average vs. willingness concerns NA replaced
#| echo: false

# using the df that replaced empty entries with "not at all" because this is conceptually correct
cor_openmindedness_mean_vs_concerns_na_replaced <- concerns_wide_na_replaced %>% map_dfr(~ tidy(cor.test(.x, openmindedness_summary_by_subject_df$mean_openmindedness)), .id = "concern") 

# view results
cor_openmindedness_mean_vs_concerns_simplified_na_replaced <- cor_openmindedness_mean_vs_concerns_na_replaced %>% 
    mutate(
    significance = case_when(
      p.value < 0.001 ~ "< 0.001",
      p.value < 0.01  ~ "< 0.01",
      p.value < 0.05  ~ "< 0.05",
      TRUE            ~ "ns"  # not significant
    ),
    direction = if_else(estimate < 0, "negative", "positive"),
    estimate_rounded = round(estimate, 3),
    pvalue_rounded = signif(p.value, 3)
  ) %>%
  select(concern, estimate_rounded, pvalue_rounded, direction, significance) %>%
  full_join(question_key_for_fa, by = c("concern" = "short_label")) %>%
    mutate(across(everything(), ~ gsub('unwillingness-reason-', '', .))) 

sig_count_cor_openmindedness_mean_vs_concerns_simplified_na_replaced <- sum(cor_openmindedness_mean_vs_concerns_simplified_na_replaced$significance != "ns")

```

```{r}
#| label: Function for printing correlation results
#| echo: false


print_significant_correlations <- function(df) {
  formatted_items <- df %>%
arrange(desc(abs(as.numeric(estimate_rounded)))) %>%
    filter(significance != "ns") %>%
    mutate(
      formatted = paste0(
       '"', question, '"',
        " ($r$ = ", estimate_rounded,
        ", $p$ = ", significance, ")"
      )
    ) %>%
    pull(formatted)
  
  n <- length(formatted_items)
  
  if (n == 0) {
    return("No significant correlations found.")
  } else if (n == 1) {
    return(formatted_items)
  } else if (n == 2) {
    return(paste(formatted_items, collapse = " and "))
  } else {
    return(paste(
      paste(formatted_items[-n], collapse = ", "),
      "and", formatted_items[n]
    ))
  }
}

```

Accounting for the concerns that participants did not endorse as "not at all" impacting participants' willingness, we found that `r sig_count_cor_openmindedness_mean_vs_concerns_simplified_na_replaced` of the 25 concern items were significantly correlated with average open-mindedness.

These significant concerns were `r print_significant_correlations(cor_openmindedness_mean_vs_concerns_simplified_na_replaced)`.

```{r}
#| label: fig-hierarchical-clusters-avg-openmindedness
#| echo: false 
#| fig-cap: Box plot of average open-mindedness about topic by cluster per hierarchical analysis. 

open_mindedness_w_clusters_hierarchical <- openmindedness_summary_by_subject_df %>%
  merge(concerns_with_cluster_df, by = "prolific_subject_id")

ggplot(open_mindedness_w_clusters_hierarchical, aes(x = factor(clusterID), y = mean_openmindedness, fill = factor(clusterID))) +
  geom_boxplot() +
  labs(x = "Participant Concern Profile (Clusters)", y = "Average Open-Mindedness (1-7)", title = "Average Open-mindedness by Participant Concern Profiles") +
  theme_minimal() + 
  scale_fill_manual(values = c("1" = "#951829", "2" = "#951829", "3" = "#951829", "4" = "#951829")) +
    guides(fill = "none")


```

@fig-hierarchical-clusters-avg-openmindedness communicates a similar pattern to @fig-hierarchical-fig-clusters-avg-willingness: Cluster 1 shows the greatest concentration of open-mindedness and cluster 4 shows relatively low concentrations of open-mindedness. Notably, cluster 2's median and lower quartile are slightly lower than cluster 4's; cluster 2 still maintains the largest range, though, which remains consistent with @fig-hierarchical-fig-clusters-avg-willingness. 

### Demographics & Willingness Relationship
```{r}
#| label: fig-sex-clusters-hierarchical
#| echo: false 
#| fig-cap: Plot of sex composition by hierarchical cluster helps us explore whether sex may be correlated with types of concern profiles.

sex_cluster_plot <- ggplot(surveys_df, aes(x = clusterID, fill = Sex)) +
  geom_bar(position = "fill") +
  labs(x = "Participant Concern Profile (Cluster)", y = "Proportion of Male and Female Participants", title = "Sex Composition by Participant Concern Profile") + 
  scale_fill_manual(values = c("Male" = "#951829", "Female" = "#63666A"))

  

ethnicity_cluster_plot <- ggplot(surveys_df, aes(x = clusterID, fill = `Ethnicity simplified`)) +
   geom_bar(position = "fill") +
   labs(x = "Participant Concern Profiles (Clusters)", y = "Proportion", title = "Ethnicities Within Participant Concern Profiles") +
  scale_fill_manual(values = c("White" = "#951829", "Other" = "#63666A", Mixed = "#ED1C24", "Black" = "#D0D0CE", "Asian" = "#F2D4D7"))

age_cluster_plot <- ggplot(
  surveys_df,
  aes(x = factor(clusterID), y = as.numeric(Age), fill = factor(clusterID))
) +
  geom_boxplot() +
  labs(
    title = "Age Distributions Within Participant Concern Profiles",
    x = "Participant Concern Profiles (Clusters)",
    y = "Age (Years)"
  ) + 
  scale_fill_manual(values = c("1" = "#951829", "2" = "#951829", "3" = "#951829", "4" = "#951829")) +
    guides(fill = "none")

sex_cluster_plot
```


```{r}
#| label: Look for sign diffs in sex btwn clusters
#| echo: false

# make a df that only has one row per subject
sig_dif_df <- surveys_df %>%
  distinct(prolific_subject_id, .keep_all = TRUE) %>%
  mutate(Age = as.numeric(Age))

sex_table <- table(sig_dif_df$clusterID, sig_dif_df$Sex)
chisq_sex <- chisq.test(sex_table)
# p value isn't significant 

```

@fig-sex-clusters-hierarchical shows some differences in sex composition by cluster, particularly in cluster 3, but a chi-square test of independence indicated that sex did not significantly differ across clusters, χ²(`r chisq_sex$parameter`, N = `r sum(sex_table)`) = `r round(chisq_sex$statistic, 2)`, *p* = `r round(chisq_sex$p.value, 3)`. Cluster 3 had a relatively small number of participants, so the chi-square test may have been underpowered. 

```{r}
#| label: fig-age-clusters-hierarchical
#| echo: false 
#| fig-cap: Plot of age distribution by hierarchical cluster helps us explore whether age may be correlated with types of concern profiles.

age_cluster_plot
```


```{r}
#| label: Look for sig diffs in age btwn clusters
#| echo: false

# do anova and nonparametric just in case 
anova_age <- aov(Age ~ factor(clusterID), data = sig_dif_df)

summary_anova_age <- summary(anova_age)

nonpara_age <- kruskal.test(Age ~ factor(clusterID), data = sig_dif_df)

#neither significant 

library(car)
#leveneTest(Age ~ factor(clusterID), data = sig_dif_df)
```

@fig-age-clusters-hierarchical also shows some variation in age distributions, but a
one-way ANOVA indicated that age did not significantly differ across clusters, F(`r summary_anova_age[[1]]$Df[1]`, `r summary_anova_age[[1]]$Df[2]`) = `r round(summary_anova_age[[1]]$"F value"[1], 2)`, p `r ifelse(summary_anova_age[[1]]$"Pr(>F)"[1] < .001, "< .001", paste0("= ", sub("^0", "", format(round(summary_anova_age[[1]]$"Pr(>F)"[1], 3), nsmall = 3))))`.  


```{r}
#| label: fig-ethnicity-clusters-hierarchical
#| echo: false 
#| fig-cap: Plot of ethnicity composition by hierarchical cluster helps us explore whether ethnicity may be correlated with types of concern profiles.

ethnicity_cluster_plot
```


```{r}
#| label: Look for sign diffs in ethnicity btwn clusters
#| echo: false
#| warning: false
#| message: false 

eth_table <- table(sig_dif_df$clusterID, sig_dif_df$`Ethnicity simplified`)

#there are too few people in the third and fourth cluster, so i am going to group minority ethnicities together. the oversimplification is not ideal but at least it's something 
eth_chisq_df <- sig_dif_df %>%
  mutate(
    ethnicity_collapsed = case_when(
      `Ethnicity simplified` %in% c("Black", "Mixed", "Other", "Asian") ~ "Minority",
      `Ethnicity simplified` == "White" ~ "White"
    )
  )

eth_simplified_table <- table(eth_chisq_df$clusterID, eth_chisq_df$ethnicity_collapsed)
chisq_ethnicity <- chisq.test(eth_simplified_table)
# not significant


# gonna look at collapsing cluster similar (by concern profile) clusters instead of race  
eth_chisq_2_df <- sig_dif_df %>%
  mutate(collapsed_clusterID = case_when(
    clusterID == 1 ~ "one_and_two",
    clusterID == 2 ~ "one_and_two",
    clusterID == 3 ~ "three_and_four",
    clusterID == 4 ~ "three_and_four",
    .default = NA
  ))

eth_simplified_table_2 <- table(eth_chisq_2_df$collapsed_clusterID, eth_chisq_2_df$`Ethnicity simplified`)
chisq_ethnicity_2 <- chisq.test(eth_simplified_table_2)
# still too few obs. could do more extensive methods, but i don't think it's worth it here

```


@fig-ethnicity-clusters-hierarchical showed some variation as well, but a chi-square test of independence indicated that ethnicity did not significantly differ across clusters, χ²(`r chisq_ethnicity$parameter`, N = `r sum(eth_simplified_table)`) = `r round(chisq_ethnicity$statistic, 2)`, *p* = `r round(chisq_ethnicity$p.value, 3)`. 

```{r}
#| label: fig-topic-choice-clusters-hierarchical
#| echo: false 
#| fig-cap: Plots of topic choice by hierarchical cluster helps us explore whether the clusters may have significant differences in which topics they focused on.

# within each cluster, finding percent of subjects on each topic. i tried to do this in one df but kept getting an error that i couldn't figure out 
hierarchical_cluster1_topic_percent <- surveys_df %>%
  filter(clusterID == 1) %>%
  distinct(prolific_subject_id, .keep_all = TRUE) %>%
  group_by(topicChoiceAsString) %>%
  summarise(n_cluster_1 = n(), .groups = "drop") %>%
  mutate(percent_cluster_1 = n_cluster_1 / sum(n_cluster_1))

hierarchical_cluster2_topic_percent <- surveys_df %>%
  filter(clusterID == 2) %>%
  distinct(prolific_subject_id, .keep_all = TRUE) %>%
  group_by(topicChoiceAsString) %>%
  summarise(n_cluster_2 = n(), .groups = "drop") %>%
  mutate(percent_cluster_2 = n_cluster_2 / sum(n_cluster_2))

hierarchical_cluster3_topic_percent <- surveys_df %>%
  filter(clusterID == 3) %>%
  distinct(prolific_subject_id, .keep_all = TRUE) %>%
  group_by(topicChoiceAsString) %>%
  summarise(n_cluster_3 = n(), .groups = "drop") %>%
  mutate(percent_cluster_3 = n_cluster_3 / sum(n_cluster_3))

hierarchical_cluster4_topic_percent <- surveys_df %>%
  filter(clusterID == 4) %>%
  distinct(prolific_subject_id, .keep_all = TRUE) %>%
  group_by(topicChoiceAsString) %>%
  summarise(n_cluster_4 = n(), .groups = "drop") %>%
  mutate(percent_cluster_4 = n_cluster_4 / sum(n_cluster_4))

#join together
hierarchical_cluster_topic_percent <- hierarchical_cluster1_topic_percent %>%
  full_join(hierarchical_cluster2_topic_percent, by = "topicChoiceAsString") %>% 
  full_join(hierarchical_cluster3_topic_percent, by = "topicChoiceAsString") %>%
 full_join(hierarchical_cluster4_topic_percent, by = "topicChoiceAsString") 

library(forcats)

surveys_df$short_topic_labels <- fct_infreq(surveys_df$short_topic_labels)

ggplot(surveys_df, aes(x = clusterID, fill = short_topic_labels)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c(
"U.S. in Israeli-Palestinian conflict" = "#641A2B",                              "Transgender athletes in sports" = "#951829",
"Human euthanasia" = "#ED1C24",
"Capitalism in economy" = "#F2D4D7",
"Climate change policies" = "#FFF8EF",
"Deportation policies" = "#D0D0CE",
"A.I. in human-run spaces" = "#808080",
"Vaccine Mandates" = "#63666A",
"Police conduct and accountability" = "#C6AA76",
"Government in healthcare" = "#000080"
)) +
  labs(x = "Participant Concern Profiles (Clusters)", y = "Proportion", title = "Focus Topic Distributions Within Participant Concern Profiles", fill = "Focus topic")

```

```{r}
#| label: Look for sign diffs in focus topic btwn clusters
#| echo: false
#| warning: false
#| message: false 

# topic choice
topic_table <- table(sig_dif_df$clusterID, sig_dif_df$short_topic_labels)
#there are too few obs in a lot of entries, so i'm going to collapse the less common topics together 

# identify rare topics dynamically
rare_topics <- topic_frequency_counts %>%
  filter(n < 30) %>% # min count topics 
  pull(short_topic_labels)

# collapse clusters 3 and 4 and collapse low count topics because there are too few counts to run chi sq test
topic_chisq_df <- sig_dif_df %>%
  mutate(collapsed_clusterID = case_when(
    clusterID == 1 ~ "one",
    clusterID == 2 ~ "two",
    clusterID == 3 ~ "three_and_four",
    clusterID == 4 ~ "three_and_four",
    .default = NA
  )) %>%
  mutate(topic_collapsed = ifelse(short_topic_labels %in% rare_topics, 
                             "Other", 
                             short_topic_labels)) 
topic_table_coll <- table(topic_chisq_df$collapsed_clusterID, topic_chisq_df$topic_collapsed)

chisq_topic_coll <- chisq.test(topic_table_coll)
#insignificant p value. the data has become so simplified anyway

# try monte carlo simulation 
topic_table_original <- table(sig_dif_df$clusterID, sig_dif_df$short_topic_labels)

# B = 10000 tells R to run 10,000 simulations to find the exact p-value
chisq_simulated_topic <- chisq.test(topic_table_original, simulate.p.value = TRUE, B = 10000)

#print(chisq_simulated_topic)
```

@fig-topic-choice-clusters-hierarchical  also showed some variation, but a Monte Carlo chi-square test of independence (based on 10,000 replicates) indicated that focus topic did not significantly differ across clusters, χ²(N = `r sum(topic_table_original)`) = `r round(chisq_simulated_topic$statistic, 2)`, *p* = `r round(chisq_simulated_topic$p.value, 3)`. I used Monte Carlo because of the low counts for some clusters in some topics, but I could have also tried other statistical significance measures. The differences look relatively random and this is exploratory anyway, so I was not too concerned with nuances in methodology here. 

### Overall Willingness & Willingness Concerns Relationships

```{r}
#| label: fig-EDA-correlation-relationships 
#| echo: false 
#| warning: false
#| message: false
#| fig-cap: Relationships between each willingness concern and average willingness to converse. Using a 7-pt Likert scale, participants rated the impact of each willingness concern on their overall willingness. Using a slider from 0 to 100, participants rated 2 statements about their willingness to converse, which we averaged to find each participant's average willingness to converse about their focus topic. 

library(gridExtra) # so we can print the plots in columns 

concerns_and_avg_willingness_df <- concerns_labeled_wide %>%
  full_join(avg_willingness_df, by = "prolific_subject_id")

# see the average number of concerns that people indicated as affecting them 

# Create all the individual plots
p1 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q01, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p2 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q02, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p3 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q03, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p4 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q04, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p5 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q05, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p6 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q06, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p7 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q07, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p8 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q08, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p9 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q09, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p10 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q010, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p11 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q011, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p12 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q012, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p13 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q013, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p14 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q014, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p15 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q015, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p16 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q016, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p17 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q017, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p18 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q018, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p19 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q019, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p20 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q020, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p21 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q021, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p22 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q022, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p23 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q023, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p24 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q024, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)
p25 <- ggplot(concerns_and_avg_willingness_df, aes(x = Q025, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0)

# arrange the plots in a 2x8 grid
#suppressWarnings(grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, p13, p14, p15, p16, p17, p18, p19, p20, p21, p22, p23, p24, p25, ncol = 2))


### the results seem to pretty much show no relationship, but if you squint your eyes, there is vaguely a linear relationship in some of the variables. i'm going to stick with pearson because, to my knowledge, the data doesn't meet the standards for any non-parametric methods 


# print an example plot
ggplot(concerns_and_avg_willingness_df, aes(x = Q04, y = avg_willingness)) + geom_point() + geom_jitter(width = 0.2, height = 0) + labs(x = "Rating of concern item's effect on willingness (1-7)", y = "Willingness to converse (0-100)", title = "Willingness to Converse vs. Severity of Concern Effect on Willingness", subtitle = "Includes ratings of all concern items")
```

```{r}
#| label: Correlations willingness slider average vs. willingness concerns (NA is not replaced) 
#| echo: false

#prepping data
reasons_and_willingness_merged_df <- merge(concerns_labeled_wide, avg_willingness_df, by = 'prolific_subject_id') %>%
  ungroup() %>%
  select(-prolific_subject_id, -topicChoiceAsString)

#run cor test
cor_willingness_slider_vs_concerns <- reasons_and_willingness_merged_df %>% select(-avg_willingness) %>%
  map_dfr( ~ tidy(cor.test(.x, reasons_and_willingness_merged_df$avg_willingness)),
  .id = "reason_item"
)

#process results 
cor_willingness_slider_vs_concerns <- cor_willingness_slider_vs_concerns %>%
  mutate(
    significance = case_when(
      p.value < 0.001 ~ "< 0.001",
      p.value < 0.01  ~ "< 0.01",
      p.value < 0.05  ~ "< 0.05",
      TRUE            ~ "ns"  # not significant
    ),
    direction = if_else(estimate < 0, "negative", "positive"),
    estimate_rounded = round(estimate, 3),
    pvalue_rounded = signif(p.value, 3)
  ) %>%
  select(reason_item, estimate_rounded, pvalue_rounded, direction, significance)

```

```{r}
#| label: Correlations willingness slider average vs. willingness concerns NA replaced 
#| echo: false

#prepping data
reasons_and_willingness_merged_df_na_replaced <- merge(concerns_labeled_wide_na_replaced, avg_willingness_df, by = 'prolific_subject_id') %>%
  ungroup() %>%
  select(-prolific_subject_id, -topicChoiceAsString)

#run cor test
cor_willingness_slider_vs_concerns_na_replaced_d <- reasons_and_willingness_merged_df_na_replaced %>% 
  select(starts_with("Q0")) %>%
  map_dfr( ~ tidy(cor.test(.x, reasons_and_willingness_merged_df_na_replaced$avg_willingness)),
  .id = "reason_item"
)

#process results   
cor_willingness_slider_vs_concerns_na_replaced <- cor_willingness_slider_vs_concerns_na_replaced_d %>%
  mutate(
    significance = case_when(
      p.value < 0.001 ~ "< 0.001",
      p.value < 0.01  ~ "< 0.01",
      p.value < 0.05  ~ "< 0.05",
      TRUE            ~ "ns"  # not significant
    ),
    direction = if_else(estimate < 0, "negative", "positive"),
    estimate_rounded = round(estimate, 3),
    pvalue_rounded = signif(p.value, 3)
  ) %>%
  select(reason_item, estimate_rounded, pvalue_rounded, direction, significance) %>%
  full_join(question_key_for_fa, by = c("reason_item" = "short_label")) %>%
    mutate(across(everything(), ~ gsub('unwillingness-reason-', '', .))) %>%
  mutate(question = short_concern_label)

sig_count_cor_willingness_slider_vs_concerns_na_replaced <- sum(cor_willingness_slider_vs_concerns_na_replaced$significance != "ns")

```

I looked for correlations between each willingness concern and average willingness as indicated by the two slider questions. This served as something like a sanity check as I would expect more severe ratings of concern items to be associated with lower willingness to converse; however, a lack of correlation could be  explained by the idea that one concern item's severity may not register strongly enough against the type of background of 24 other concern items with which it is typically associated to singularly be correlated with willingness to converse. 

More than half of the concern items were negatively associated with willingness, suggesting that, for most items, a greater impact of the concern on an American's willingness is associated with lower willingness to engage.

These `r sig_count_cor_willingness_slider_vs_concerns_na_replaced` concern items were `r print_significant_correlations(cor_willingness_slider_vs_concerns_na_replaced)`.

## Discussion

### Comprehensiveness of Concerns

Through a qualitative analysis of participants' free-response explanations for avoiding difficult conversations, we found that the concerns participants raised were well-represented in the 25-item inventory. The few concerns that did not map cleanly onto existing items appeared as one-off mentions rather than true gaps. This suggests that the questionnaire comprehensively covers the primary reasons Americans report for unwillingness to engage in conversations on controversial topics, and that further item additions are not warranted at this stage. It is worth noting here, though, that low-concern profiles dominate the sample, which does not align with the intuitive expectation that we would see a more dramatic trend of conversational reluctance amid the rise of American affective polarization (e.g., @boxell2020). 

### Structure of Concerns

Our second objective was to explore the latent structure of the 25 concerns. An Exploratory Factor Analysis suggested a four-factor solution accounting for `r total_var_explained_percent`% of the total variance: Social & Reputation Consequences, Perceived Futility & Emotional Cost, Unpreparedness for Complex Topic, and Personal Vulnerability & Identity Threat. PCA on the same items recovered some related structures: PC2 mapped onto Perceived Futility & Emotional Cost, PC3 contrasted Unpreparedness against Social Consequences, and PC4 contrasted topic unfamiliarity with anticipated negative affect, which supports the robustness of the factor structure across methods. The clustering analysis, however, recovered only one factor-aligned profile (Cluster 2, elevated on futility/emotional cost). The more selective Clusters 3 and 4 instead combined items across factors, suggesting that participants tend to experience these concerns in combinations rather than as cleanly siloed dimensions. The factor structure may therefore be more useful as a measurement framework than as a typology of participants, which aligns with the theoretical background for factor analysis compared to participant-centered profiling methods. Additionally, it is worth noting here that none of these methods accounted for a good majority of variance, meaning there may be more at play, that concern profiles are more uniquely individual, or that other profiling analysis methods should be employed. 


### Low-valenced ratings

A primary takeaway from this study is that participants frequently rated their willingness as higher than we had hoped for (in the sense that we want to target and measure low-willingness situations or individuals) and rated the severity of their concerns very low. This was reflected in participants' ratings of their focus topics, in their ratings of overall willingness, and in their ratings of concern item severities. Three plausible explanations come to mind to explain this outcome.

First, participants may be responding aspirationally, thus inflating results. Participants may consciously or unconsciously be reluctant to admit to the survey, themself, or both that they are moderately to strongly unwilling to converse or are swayed by certain items. 

Second, the questions may not have elicited the specific kind of reluctance we aim to measure. The topic-comfort Likert item asks about discussing the topic in a group scenario, the slider items specify a particular discussion partner (someone who disagrees, someone who pushes back), and the concern items ask about a serious conversation. Participants may have envisioned different conversational contexts across the formats and even across the concern items, and they may have not envisioned the scenarios in which they are least willing to engage. Anticipation of contextual influence partly motivated us to include these different framings, as capturing a few framings would be better than doubling down on one single framing, but this also indicates the limitation of trying to capture a very situation-dependent phenomenon in a decontextualized way. If willingness is fundamentally contextual, then asking about it abstractly may produce a noisy average rather than a stable trait-like estimate.

Third, samples for this kind of study may self-select for people who are more willing to discuss controversial topics. Prolific users who are less willing to discuss controversial topics may have skipped over this study when browsing their options more often than users who are more willing.

### Moving Forward 

#### Measuring Willingness Concerns

One direction worth exploring is asking participants to anchor their responses to a specific context, such as the context in which they are least willing to converse. We initially considered and rejected a context-specific approach of any kind on the reasoning that the intervention aims to increase willingness broadly, and we rejected the worst-case-scenario context approach because these cases may be so extreme that our intervention will likely not be radically transformative enough to evoke meaningful changes in these contexts. However, there could be merit in asking participants to consider a situation in which they recognize that engagement would be useful or beneficial but nonetheless feel unwilling. 

Given the indications above that willingness ratings are weakly valenced and likely situational, a context-anchored design may produce more interpretable ratings and a more tractable target for intervention. The argument that people will not budge on the most extreme contexts but may budge in moderate ones remains relevant: someone who will never discuss deportation policy with hostile colleagues may still meaningfully shift in their willingness to raise the topic with friends. Asking participants to nominate a context in which they feel unwilling to converse but that engagement would be beneficial or worthwhile could capture exactly the margin our intervention is more suited to move. The tradeoff here, though, would be placing the responsibility of situation identification on the participants: whether participants can readily generate such a context, and how many will feel that one exists for them, remain open questions.

#### Applying Findings to MOCA 

##### For Assessing MOCA's Effectiveness

I see two main ways to apply these findings to assess MOCA's effectiveness. Currently, the experiment structure for assessing MOCA's effectiveness looks like having participants complete a survey, which includes the very first rudimentary draft of this questionnaire, both before and after conversing with MOCA. This allows us to assess whether MOCA ameliorates any issues in qualities like open-mindedness or in willingness concerns.  

One option is to implement the full 25-item list in the MOCA experiment's survey to gain extensive insight into MOCA's strengths and weaknesses. This would be very informative and interesting to analyze; however, it could give rise to participant fatigue and would increase study duration (which would in turn require more funding or fewer participants to conduct).

If a 25-item list is too extensive, we could instead draw on factor analysis. This could look like summarizing the items within each factor into a few sentences and using these, along with the factor titles, as factor descriptors. Participants could then rate the severity of each factor's effect on their willingness to converse, reducing survey length while still capturing useful information. This would be much less informative compared to the full questionnaire and would potentially introduce noise that could cause our evaluations of MOCA's effectiveness in each factor to converge toward neutrality; however, this method could also pan out to be very informative, and it is much more resource efficient.


##### For Improving MOCA

In an ideal conversation with MOCA, the human would accurately report their main concerns to MOCA (which MOCA asks about in one of the initial messages) and then MOCA's subsequent behaviors would in part be shaped to address these concerns (which MOCA is currently prompted to do). This would allow MOCA to be flexibly shaped on a case-by-case basis in real-world usage to address each individual's concerns. 

In an experimental setting, the concern items that an individual endorses and the severity of these concerns could be reported to MOCA for even more customized intervention; however, this would require the same questionnaire procedure in real-world applications which may not be very realistic. 

Currently, MOCA addresses a user's concerns through a prompt that simply tells MOCA to identify the user's concerns (which should be explicitly stated by the user in response to MOCA's question of why the user is reluctant) and then respond with these concerns in mind. Perhaps, though, providing MOCA with this study's questionnaire and results could allow MOCA to perform this identification phase with the background information that would allow it to identify the fundamental concerns that a user has. This could then improve MOCA's ability to target the user's concerns. For example, a user might say they do not want to converse about a topic because it stresses them out. MOCA would then be able to draw on the 25-item inventory to identify the candidates that might underlie the umbrella of stress, and then give candidate-informed responses. This could potentially allow the conversation to better address the user's true concerns. Additionally, perhaps MOCA could perform a second  identification step partway through the conversation; in this phase, MOCA could refine the list of candidate items by assessing which candidates seem to be surfacing in the user's messages. 

In the longer term, were MOCA developed into an agent that could learn from its conversations with users (especially if something like a quick rate-your-experience button were added to the ends of conversations), MOCA would potentially be able to develop a skill set for helping users based on their concern profile or even on each of their concerns. I am not sure how this would work while using a proprietary LLM and needing to maintain user privacy, but this could be very influential in MOCA's performance as an intervention.

### Limitations

The primary limitation of this study is the reliance on self-report measures for all key variables (concerns, willingness, and open-mindedness). I've offered some explanations for the within-subject differences in ratings of overall willingness and for the relatively low-valenced ratings amidst the current climate of American discourse, but these results may also reflect the inherent difficulties in measuring these constructs. Future work should integrate behavioral or implicit measures to provide a more objective assessment.

In conclusion, this pretest supports the 25-item questionnaire as a valuable tool for measuring the multifaceted barriers to productive discourse on controversial topics. 

## References

